---
title: "corr_gender"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

This file calculates, visualises and summarises the correlational analyses from  the individual differences face familiarity study with controlled influence of Gender on the CCMT scores.

```{r}
#| label: setup
#| include: false

pkgs <- c(
  "EMC2", "tidyverse", "tidybayes", "patchwork", "paletteer", "here",
  "corrplot", "ggdist", "RColorBrewer", "parallel", "future", "brms"
  )
vapply(
  pkgs, library, logical(1), character.only = TRUE, logical.return = TRUE
  )

theme_set(
  theme_bw() +
    theme(
      text  = element_text(size = 15, face = "bold"),
      title = element_text(size = 15, face = "bold"),
      legend.position = "none"
    )
)

## set options
options(mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()

detectCores()

plan(multicore)

## make things reproducible where possible 
set.seed(123)

safe_pal   <- paletteer_d("rcartocolor::Safe", 4)
mycolors <- c("novel" = safe_pal[1], "learned" = safe_pal[2])
```


```{r}
#| label: read-cambridge-data

## read in the pid data
cambridge_acc_pid <- read_csv(
  here("data", "cambridge_acc_scores_pid.csv")
)
head(cambridge_acc_pid)
glimpse(cambridge_acc_pid)

## filter and wrangle
## face scores
cam_face <- cambridge_acc_pid |>
  filter(task == "face") |>
  rename(face_recog = n_correct) |>
  mutate(
    face_recog_centred = face_recog - mean(face_recog, na.rm = TRUE)
  ) |>
  select(pid, face_recog, face_recog_centred)

## car scores
cam_car <- cambridge_acc_pid |>
  filter(task == "car") |>
  rename(car_recog = n_correct) |>
  mutate(
    car_recog_centred = car_recog - mean(car_recog, na.rm = TRUE)
  ) |>
  select(pid, Sex, car_recog, car_recog_centred)
```
# 1. Regression model to control for gender effect on CCMT ##
## set defaults and create a function

```{r}
#| label: brms-defaults-function

brm_defaults <- list(
  family = gaussian(),
  iter = 2000,
  warmup = 1000,
  cores = 28,
  chains = 4,
  save_pars = save_pars(all = TRUE),
  seed = 123,
  control = list(adapt_delta = 0.95, max_treedepth = 15)
)

fit_model <- function(formula, prior, file, data, ...) {
  brm(
    formula = formula,
    prior = prior,
    file = file,
    data = data,  
    family = brm_defaults$family,
    iter = brm_defaults$iter,
    warmup = brm_defaults$warmup,
    cores = brm_defaults$cores,
    chains = brm_defaults$chains,
    save_pars = brm_defaults$save_pars,
    seed = brm_defaults$seed,
    control = brm_defaults$control,
    ...
  )
}
```

## fit model
```{r}
#| label: fit-m1-gender

## formula
formula = bf(car_recog_centred ~ 0 + Sex)

## check priors
get_prior(formula, data = cam_car, family = gaussian())

## set priors
priors <- c(
  set_prior("normal(3, 1)", class = "b"),
  set_prior("normal(0, 1)", class = "sigma")
)

## fit the model
m1_gender <- fit_model(
  formula = formula,
  prior = priors,
  data = cam_car,
  file = "models/m1_gender",
)

summary(m1_gender)

#get the residuals of the model
CCMT_residuals <- as.data.frame(residuals(m1_gender))
```

## Add residuals to the data
```{r}
#add residulas to the data used for future analysis
cam_car <- cam_car |>
  mutate(CCMT_residuals = CCMT_residuals$Estimate)

## join together for cor plots
cam_face_car <- cam_face |> 
  left_join(cam_car, by = "pid") |>
  select(pid, face_recog_centred, CCMT_residuals) |>
  rename(car = "CCMT_residuals", face = "face_recog_centred")
head(cam_face_car)

write_csv(cam_face_car, here("data", "cambridge_acc_scores_pid_res.csv"))
```

# 2. Read in posterior params

```{r}
#| label: read-post

theta_pid <- read_csv(
  "data/theta_pid.csv",
  show_col_types = FALSE
)

theta_pid_q <- read_csv(
  "data/theta_pid_q.csv",
  show_col_types = FALSE
)
```

# 3. Plot Cambridge test data

## plot face by car corr dot plot ##

```{r}
#| label: plot-face-car-corr

## face by car cor
# compute r and CI
cor_test <- cor.test(cam_face_car$face_recog_centred, cam_face_car$CCMT_residuals)
r_label  <- sprintf("r = %.2f [%.2f, %.2f]",
                    cor_test$estimate,
                    cor_test$conf.int[1],
                    cor_test$conf.int[2])

# base plot
p_face_car <- ggplot(cam_face_car, aes(x = face_recog_centred, y = CCMT_residuals)) +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "lm", colour = "blue", fill = "lightblue") +
  annotate("text",
           x    = -Inf, y = Inf,
           hjust = -0.1, vjust = 1.5,
           label = r_label,
           size  = 5, fontface = "bold") +
  labs(
    x = "Face recognition score centred",
    y = "Residuals of Car recognition score",
    title = "Cambridge face vs. car memory scores"
  )

# add marginal densities
p_face_car_dens <- ggExtra::ggMarginal(
  p_face_car,
  type     = "density",
  fill     = "steelblue",
  alpha    = 0.4,
  colour   = "steelblue"
)
p_face_car_dens

png(here("figures", "corr", "face_car-res_corr.png"), 
    width = 8, height = 6, units = "in", res = 300)
print(p_face_car_dens)
dev.off()
```

# 4. Plot parameter estimates as a function of Cambriddge test scores

## all params

wrangle

```{r}
#| label: wrangle-correlations

# join median params with cambridge scores
cor_data <- theta_pid_q |>
  select(pid, param, value) |>
  pivot_wider(names_from = param, values_from = value) |>
  left_join(cam_face, by = "pid") |>
  left_join(cam_car,  by = "pid") |>
  # derived contrasts
  mutate(
    # drift discrimination (true - false) per condition
    discrim_novel   = v_Snovel_lMTRUE   - v_Snovel_lMFALSE,
    discrim_learned = v_Slearned_lMTRUE - v_Slearned_lMFALSE,
    # cross-condition drift contrast (novel - learned discrimination)
    discrim_diff    = discrim_novel - discrim_learned,
    # threshold contrast
    thresh_diff     = B_lRnovel - B_lRlearned
  )
```

plot

```{r}
#| label: plot-raw-params

param_cols <- c(
  "v_Snovel_lMTRUE", "v_Snovel_lMFALSE",
  "v_Slearned_lMTRUE", "v_Slearned_lMFALSE",
  "sv_lMTRUE", "B_lRnovel", "B_lRlearned", "A", "t0"
)

# pivot to long with both predictors as a variable
cor_long <- cor_data |>
  pivot_longer(
    cols      = all_of(param_cols),
    names_to  = "param",
    values_to = "estimate"
  ) |>
  pivot_longer(
    cols      = c(face_recog_centred, CCMT_residuals),
    names_to  = "predictor",
    values_to = "score"
  ) |>
  mutate(
    predictor = if_else(str_detect(predictor, "face"), "Face", "Car"),
    predictor = factor(predictor, levels = c("Face", "Car"))
  )

# compute r per param x predictor
cor_labels <- cor_long |>
  group_by(param, predictor) |>
  summarise(
    r = cor(estimate, score, use = "complete.obs"),
    .groups = "drop"
  ) |>
  mutate(
    label = sprintf("r = %.2f", r),
    vjust = if_else(predictor == "Face", 1.5, 3.0)
  )

# plot
p_raw_params <- ggplot(cor_long, aes(x = score, y = estimate,
                                      colour = predictor, fill = predictor)) +
  geom_point(alpha = 0.3, size = 2) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.3) +
  geom_text(
    data = cor_labels,
    aes(label = label, vjust = vjust, colour = predictor),
    x = -Inf, y = Inf,
    hjust = -0.1,
    size = 5, fontface = "bold",
    show.legend = FALSE
  ) +
  facet_wrap(~param, scales = "free_y", ncol = 3) +
  scale_colour_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  scale_fill_manual(values   = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  labs(
    title   = "Raw parameter estimates vs. recognition scores",
    x       = "Recognition score (centred)",
    y       = "Parameter estimate (median)",
    colour  = NULL, fill = NULL
  ) +
  theme(legend.position = "bottom")

p_raw_params

ggsave(here("figures", "corr", "all_params-res.jpeg"), 
       width = 15, height = 8)
```

```{r}
#| label: plot-all-params-grid

# all params + derived contrasts in one long df
all_params_long <- cor_data |>
  select(pid, face_recog_centred, car_recog_centred,
         all_of(param_cols), 
         discrim_novel, discrim_learned, discrim_diff, thresh_diff) |>
  pivot_longer(
    cols      = -c(pid, face_recog_centred, CCMT_residuals),
    names_to  = "param",
    values_to = "estimate"
  )

# r labels per param x predictor
all_r_face <- all_params_long |>
  group_by(param) |>
  summarise(
    r     = cor(estimate, face_recog_centred, use = "complete.obs"),
    .groups = "drop"
  ) |>
  mutate(label = sprintf("%s\n(r = %.2f)", param, r))

all_r_car <- all_params_long |>
  group_by(param) |>
  summarise(
    r     = cor(estimate, CCMT_residuals, use = "complete.obs"),
    .groups = "drop"
  ) |> 
  mutate(label = sprintf("%s\n(r = %.2f)", param, r))

# join labels back
face_plot_data <- all_params_long |>
  left_join(all_r_face, by = "param") |>
  mutate(param_label = label)

car_plot_data <- all_params_long |>
  left_join(all_r_car, by = "param") |>
  mutate(param_label = label)

# face plot
p_all_face <- ggplot(face_plot_data,
                      aes(x = face_recog_centred, y = estimate)) +
  geom_point(alpha = 0.4, size = 2, colour = "grey30") +
  geom_smooth(method = "lm", se = TRUE, colour = "red", fill = "grey30") +
  facet_wrap(~param_label, scales = "free_y", ncol = 5) +
  labs(
    title = "All parameters: correlations with face recognition",
    x     = "Face recognition ability (centred)",
    y     = "Parameter value"
  )

p_all_face

ggsave(here("figures", "corr", "all_params_face.jpeg"), 
       width = 15, height = 8)

# car plot
p_all_car <- ggplot(car_plot_data,
                     aes(x = CCMT_residuals, y = estimate)) +
  geom_point(alpha = 0.4, size = 2, colour = "grey30") +
  geom_smooth(method = "lm", se = TRUE, colour = "red", fill = "grey30") +
  facet_wrap(~param_label, scales = "free_y", ncol = 5) +
  labs(
    title = "All parameters: correlations with car recognition",
    x     = "Car recognition ability (centred)",
    y     = "Parameter value"
  )

p_all_car

ggsave(here("figures", "corr", "all_params_car-res.jpeg"), 
       width = 15, height = 8)
```

# 6. Plausible value correlations

```{r}
#| label: plausible-value-wrangle

all_prereg_params <- c(
  param_cols,
  "discrim_novel", "discrim_learned",
  "discrim_diff", "thresh_diff"
)

param_labels <- c(
  B_lRnovel        = "Threshold: Novel",
  B_lRlearned      = "Threshold: Learned",
  thresh_diff      = "Threshold: Novel - Learned",
  v_Snovel_lMTRUE  = "Drift: Novel (true)",
  v_Snovel_lMFALSE = "Drift: Novel (false)",
  discrim_novel    = "Drift: Novel (true - false)",
  v_Slearned_lMTRUE  = "Drift: Learned (true)",
  v_Slearned_lMFALSE = "Drift: Learned (false)",
  discrim_learned  = "Drift: Learned (true - false)",
  discrim_diff     = "Drift: (Novel - Learned) difference"
)

# join theta_pid (full posterior) with cambridge scores
# theta_pid has 300 pids x 9 params x 15000 draws
pv_data <- theta_pid |>
  pivot_wider(names_from = param, values_from = value) |>
  mutate(
    discrim_novel   = v_Snovel_lMTRUE - v_Snovel_lMFALSE,
    discrim_learned = v_Slearned_lMTRUE - v_Slearned_lMFALSE,
    discrim_diff    = discrim_novel - discrim_learned,
    thresh_diff     = B_lRnovel - B_lRlearned
  ) |>
  left_join(
    cam_face |> select(pid, face_recog_centred),
    by = "pid"
  ) |>
  left_join(
    cam_car |> select(pid, CCMT_residuals),
    by = "pid"
  ) |>
  pivot_longer(
    cols      = all_of(all_prereg_params),
    names_to  = "param",
    values_to = "estimate"
  )

# compute correlation for each draw x param x predictor
# this will take a moment with 15000 draws
pv_cors <- pv_data |>
  group_by(.draw, param) |>
  summarise(
    r_face = cor(estimate, face_recog_centred, use = "complete.obs"),
    r_car = cor(estimate, CCMT_residuals, use = "complete.obs"),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols      = c(r_face, r_car),
    names_to  = "predictor",
    values_to = "r"
  ) |>
  mutate(
    predictor = if_else(predictor == "r_face", "Face", "Car"),
    predictor = factor(predictor, levels = c("Face", "Car"))
  )

# save out — this is a big object
write_csv(pv_cors, here("data", "pv_cors-res.csv"))
```

```{r}
#| label: plausible-value-summary

# summarise
pv_summary <- pv_cors |>
  group_by(param, predictor) |>
  median_qi(r, .width = c(0.95)) |>
  mutate(label = param_labels[param]) |> 
  print(n=Inf)
```

```{r}
#| label: plot-pv-key-params
#| fig-width: 12
#| fig-height: 8

# key params per prereg: drift discriminations + threshold + their differences
key_params <- c(
  "discrim_novel", "discrim_learned", "discrim_diff",
  "B_lRnovel", "B_lRlearned", "thresh_diff"
)

key_labels <- c(
  discrim_novel   = "Drift: Novel (true-false)",
  discrim_learned = "Drift: Learned (true-false)",
  discrim_diff    = "Drift: Novel - Learned",
  B_lRnovel       = "Threshold: Novel",
  B_lRlearned     = "Threshold: Learned",
  thresh_diff     = "Threshold: Novel - Learned"
)

pv_cors_key <- pv_cors |>
  filter(param %in% key_params) |>
  mutate(
    param = factor(param, levels = key_params,
                   labels = key_labels),
    predictor = factor(predictor, levels = c("Car", "Face"))
  )

ggplot(pv_cors_key, aes(x = r, y = predictor, 
                         fill = predictor, colour = predictor)) +
  stat_halfeye(alpha = 0.7, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  scale_colour_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  facet_wrap(~param, ncol = 3) +
  labs(
    title = "Posterior correlation distributions — key parameters",
    x     = "Plausible correlation with recognition score",
    y     = NULL,
    fill  = NULL, colour = NULL
  ) +
  theme(legend.position = "bottom")

ggsave(here("figures", "pv_cors_key-res.jpeg"), width = 12, height = 8)
```