---
title: "Correlational analyses"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

This file calculates, visualises and summarises the correlational analyses from  the individual differences face familiarity study.

```{r}
#| label: setup
#| include: false

pkgs <- c(
  "EMC2", "tidyverse", "tidybayes", "patchwork", "paletteer", "here",
  "corrplot", "ggdist", "RColorBrewer", "parallel", "future"
  )
vapply(
  pkgs, library, logical(1), character.only = TRUE, logical.return = TRUE
  )

theme_set(
  theme_bw() +
    theme(
      text  = element_text(size = 15, face = "bold"),
      title = element_text(size = 15, face = "bold"),
      legend.position = "none"
    )
)

## set options
options(mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()

detectCores()

plan(multicore)

## make things reproducible where possible 
set.seed(123)

safe_pal   <- paletteer_d("rcartocolor::Safe", 4)
mycolors <- c("novel" = safe_pal[1], "learned" = safe_pal[2])
```

# 1. Read in the raw Cambridge test data and wrangle

raw Cambridge test data

```{r}
#| label: read-cambridge-data

## read in the pid data
cambridge_acc_pid <- read_csv(
  here("data", "cambridge_acc_scores_pid.csv")
)
head(cambridge_acc_pid)
glimpse(cambridge_acc_pid)

## filter and wrangle
## face scores
cam_face <- cambridge_acc_pid |>
  filter(task == "face") |>
  rename(face_recog = n_correct) |>
  mutate(
    face_recog_centred = face_recog - mean(face_recog, na.rm = TRUE)
  ) |>
  select(pid, face_recog, face_recog_centred)

## car scores
cam_car <- cambridge_acc_pid |>
  filter(task == "car") |>
  rename(car_recog = n_correct) |>
  mutate(
    car_recog_centred = car_recog - mean(car_recog, na.rm = TRUE)
  ) |>
  select(pid, car_recog, car_recog_centred)

## join together for cor plots
cam_face_car <- cam_face |> 
  left_join(cam_car, by = "pid")
head(cam_face_car)
```

# 2. Read in the model object and wrangle

```{r}
#| label: read-model
#| eval: false 

LBA_iid <- readRDS("models/LBA_iid.rds")
```

```{r}
#| label: wrangle-post
#| eval: false ## just do this once

## col names for map = TRUE
param_cols <- c(
  "v_Snovel_lMTRUE", "v_Snovel_lMFALSE",
  "v_Slearned_lMTRUE", "v_Slearned_lMFALSE",
  "sv_lMTRUE", "B_lRnovel", "B_lRlearned", "A", "t0"
)

# subject-level — alpha, natural scale, drop constant
params <- EMC2::parameters(
  LBA_iid,
  selection = "alpha",
  N = NULL,
  map = TRUE
) |>
  select(-sv_lMFALSE)

# take a look
head(params)
glimpse(params)

# pid-level posteriors
n_pids  <- length(unique(params$subjects))
n_draws <- nrow(params) / n_pids

theta_pid <- params |>
  as_tibble() |>
  pivot_longer(
    cols = all_of(param_cols),
    names_to = "param",
    values_to = "value"
  ) |>
  rename(pid = subjects) |>
  mutate(.draw = rep(1:n_draws, each = length(param_cols), times = n_pids))

head(theta_pid)
glimpse(theta_pid)

## pid summary data
## 300 pids * 9 params = 2700
theta_pid_q <- theta_pid |>
  group_by(pid, param) |>
  median_qi(value)

## save out some files
write_csv(theta_pid, "data/theta_pid.csv")
write_csv(theta_pid_q, "data/theta_pid_q.csv")
```

read in posterior params

```{r}
#| label: read-post

theta_pid <- read_csv(
  "data/theta_pid.csv",
  show_col_types = FALSE
)

theta_pid_q <- read_csv(
  "data/theta_pid_q.csv",
  show_col_types = FALSE
)
```

# 3. Plot Cambridge test data

## plot face recog distribution ##

face recog data

```{r}
#| label: plot-face-car-hist

p3.1 <- cam_face |>
  ggplot(aes(x = face_recog)) +
  geom_histogram(
    bins = 20,
    fill = "skyblue",
    alpha = 0.7,
    color = "black"
  ) +
  geom_vline(
    xintercept = mean(cam_face$face_recog, na.rm = TRUE),
    linetype = "dashed",
    color = "red"
  ) +
  labs(
    title = "Face Recognition Score Distribution",
    x = "Face Recognition Score (0-54)",
    y = "Count",
    caption = "Red line = mean"
  )

p3.2 <- cam_car |>
  ggplot(aes(x = car_recog)) +
  geom_histogram(
    bins  = 20,
    fill  = "green",
    alpha = 0.7,
    color = "black"
  ) +
  geom_vline(
    xintercept = mean(cam_car$car_recog, na.rm = TRUE),
    linetype = "dashed",
    color = "red"
  ) +
  labs(
    title = "Car Recognition Score Distribution",
    x     = "Car Recognition Score (0-54)",
    y     = "Count",
    caption = "Red line = mean"
  )

(p3.1 / p3.2) +
  plot_layout(axes = "collect")
```

## plot face by car corr dot plot ##

```{r}
#| label: plot-face-car-corr

## face by car cor
# compute r and CI
cor_test <- cor.test(cam_face_car$face_recog, cam_face_car$car_recog)
r_label  <- sprintf("r = %.2f [%.2f, %.2f]",
                    cor_test$estimate,
                    cor_test$conf.int[1],
                    cor_test$conf.int[2])

# base plot
p_face_car <- ggplot(cam_face_car, aes(x = face_recog, y = car_recog)) +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "lm", colour = "blue", fill = "lightblue") +
  annotate("text",
           x    = -Inf, y = Inf,
           hjust = -0.1, vjust = 1.5,
           label = r_label,
           size  = 5, fontface = "bold") +
  labs(
    x = "Face recognition score",
    y = "Car recognition score",
    title = "Cambridge face vs. car memory scores"
  )

# add marginal densities
p_face_car_dens <- ggExtra::ggMarginal(
  p_face_car,
  type     = "density",
  fill     = "steelblue",
  alpha    = 0.4,
  colour   = "steelblue"
)
p_face_car_dens

png(here("figures", "corr", "face_car_corr.png"), 
    width = 8, height = 6, units = "in", res = 300)
print(p_face_car_dens)
dev.off()
```

# 4. Plot parameter estimates as a function of Cambriddge test scores

## all params

wrangle

```{r}
#| label: wrangle-correlations

# join median params with cambridge scores
cor_data <- theta_pid_q |>
  select(pid, param, value) |>
  pivot_wider(names_from = param, values_from = value) |>
  left_join(cam_face, by = "pid") |>
  left_join(cam_car,  by = "pid") |>
  # derived contrasts
  mutate(
    # drift discrimination (true - false) per condition
    discrim_novel   = v_Snovel_lMTRUE   - v_Snovel_lMFALSE,
    discrim_learned = v_Slearned_lMTRUE - v_Slearned_lMFALSE,
    # cross-condition drift contrast (novel - learned discrimination)
    discrim_diff    = discrim_novel - discrim_learned,
    # threshold contrast
    thresh_diff     = B_lRnovel - B_lRlearned
  )
```

plot

```{r}
#| label: plot-raw-params

param_cols <- c(
  "v_Snovel_lMTRUE", "v_Snovel_lMFALSE",
  "v_Slearned_lMTRUE", "v_Slearned_lMFALSE",
  "sv_lMTRUE", "B_lRnovel", "B_lRlearned", "A", "t0"
)

# pivot to long with both predictors as a variable
cor_long <- cor_data |>
  pivot_longer(
    cols      = all_of(param_cols),
    names_to  = "param",
    values_to = "estimate"
  ) |>
  pivot_longer(
    cols      = c(face_recog_centred, car_recog_centred),
    names_to  = "predictor",
    values_to = "score"
  ) |>
  mutate(
    predictor = if_else(str_detect(predictor, "face"), "Face", "Car"),
    predictor = factor(predictor, levels = c("Face", "Car"))
  )

# compute r per param x predictor
cor_labels <- cor_long |>
  group_by(param, predictor) |>
  summarise(
    r = cor(estimate, score, use = "complete.obs"),
    .groups = "drop"
  ) |>
  mutate(
    label = sprintf("r = %.2f", r),
    vjust = if_else(predictor == "Face", 1.5, 3.0)
  )

# plot
p_raw_params <- ggplot(cor_long, aes(x = score, y = estimate,
                                      colour = predictor, fill = predictor)) +
  geom_point(alpha = 0.3, size = 2) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.3) +
  geom_text(
    data = cor_labels,
    aes(label = label, vjust = vjust, colour = predictor),
    x = -Inf, y = Inf,
    hjust = -0.1,
    size = 5, fontface = "bold",
    show.legend = FALSE
  ) +
  facet_wrap(~param, scales = "free_y", ncol = 3) +
  scale_colour_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  scale_fill_manual(values   = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  labs(
    title   = "Raw parameter estimates vs. recognition scores",
    x       = "Recognition score (centred)",
    y       = "Parameter estimate (median)",
    colour  = NULL, fill = NULL
  ) +
  theme(legend.position = "bottom")

p_raw_params

ggsave(here("figures", "corr", "all_params.jpeg"), 
       width = 15, height = 8)
```

### a different layout plus derived corrs
 
```{r}
#| label: plot-all-params-grid

# all params + derived contrasts in one long df
all_params_long <- cor_data |>
  select(pid, face_recog_centred, car_recog_centred,
         all_of(param_cols), 
         discrim_novel, discrim_learned, discrim_diff, thresh_diff) |>
  pivot_longer(
    cols      = -c(pid, face_recog_centred, car_recog_centred),
    names_to  = "param",
    values_to = "estimate"
  )

# r labels per param x predictor
all_r_face <- all_params_long |>
  group_by(param) |>
  summarise(
    r     = cor(estimate, face_recog_centred, use = "complete.obs"),
    .groups = "drop"
  ) |>
  mutate(label = sprintf("%s\n(r = %.2f)", param, r))

all_r_car <- all_params_long |>
  group_by(param) |>
  summarise(
    r     = cor(estimate, car_recog_centred, use = "complete.obs"),
    .groups = "drop"
  ) |> 
  mutate(label = sprintf("%s\n(r = %.2f)", param, r))

# join labels back
face_plot_data <- all_params_long |>
  left_join(all_r_face, by = "param") |>
  mutate(param_label = label)

car_plot_data <- all_params_long |>
  left_join(all_r_car, by = "param") |>
  mutate(param_label = label)

# face plot
p_all_face <- ggplot(face_plot_data,
                      aes(x = face_recog_centred, y = estimate)) +
  geom_point(alpha = 0.4, size = 2, colour = "grey30") +
  geom_smooth(method = "lm", se = TRUE, colour = "red", fill = "grey30") +
  facet_wrap(~param_label, scales = "free_y", ncol = 5) +
  labs(
    title = "All parameters: correlations with face recognition",
    x     = "Face recognition ability (centred)",
    y     = "Parameter value"
  )

p_all_face

ggsave(here("figures", "corr", "all_params_face.jpeg"), 
       width = 15, height = 8)

# car plot
p_all_car <- ggplot(car_plot_data,
                     aes(x = car_recog_centred, y = estimate)) +
  geom_point(alpha = 0.4, size = 2, colour = "grey30") +
  geom_smooth(method = "lm", se = TRUE, colour = "red", fill = "grey30") +
  facet_wrap(~param_label, scales = "free_y", ncol = 5) +
  labs(
    title = "All parameters: correlations with car recognition",
    x     = "Car recognition ability (centred)",
    y     = "Parameter value"
  )

p_all_car

ggsave(here("figures", "corr", "all_params_car.jpeg"), 
       width = 15, height = 8)


#Meaningful correlations
# face plot
discrim_learned <- face_plot_data |>
  filter(param == c("discrim_learned", "discrim_novel"))
p_drift_face <- ggplot(discrim_learned,
                      aes(x = face_recog_centred, y = estimate)) +
  geom_point(alpha = 0.4, size = 2, colour = "grey30") +
  geom_smooth(method = "lm", se = TRUE, colour = "red", fill = "grey30") +
  facet_wrap(~param_label, scales = "free_y", ncol = 2) +
  labs(
    title = "Drift rate (TRUE-FALSE) correlations with face recognition",
    x     = "Face recognition ability (centred)",
    y     = "Drift rate"
  )

p_drift_face
ggsave(here("figures", "corr", "drift_face.jpeg"), 
       width = 8, height = 6)

# car plot
discrim_learned <- car_plot_data |>
  filter(param == c("discrim_learned", "discrim_novel"))
p_drift_car <- ggplot(discrim_learned,
                     aes(x = car_recog_centred, y = estimate)) +
  geom_point(alpha = 0.4, size = 2, colour = "grey30") +
  geom_smooth(method = "lm", se = TRUE, colour = "red", fill = "grey30") +
  facet_wrap(~param_label, scales = "free_y", ncol = 5) +
  labs(
    title = "Drift rate (TRUE-FALSE) correlations with car recognition",
    x     = "Car recognition ability (centred)",
    y     = "Drift rate"
  )

p_drift_car

ggsave(here("figures", "corr", "drift_car.jpeg"), 
       width = 8, height = 6)

```

## plot pre-registered relationships

wrangle

```{r}
#| label: wrangle-focused-plots

# drift long format with median estimates
drift_plot_data <- cor_data |>
  select(pid, face_recog_centred, car_recog_centred,
         v_Snovel_lMTRUE, v_Snovel_lMFALSE,
         v_Slearned_lMTRUE, v_Slearned_lMFALSE) |>
  pivot_longer(
    cols      = starts_with("v_"),
    names_to  = "param",
    values_to = "estimate"
  ) |>
  mutate(
    condition = if_else(str_detect(param, "novel"), "Novel", "Learned"),
    accuracy  = if_else(str_detect(param, "TRUE"),  "true",  "false"),
    condition = factor(condition, levels = c("Learned", "Novel")),
    accuracy  = factor(accuracy,  levels = c("true", "false"))
  )

# threshold long format
thresh_plot_data <- cor_data |>
  select(pid, face_recog_centred, car_recog_centred,
         B_lRnovel, B_lRlearned) |>
  pivot_longer(
    cols      = starts_with("B_"),
    names_to  = "param",
    values_to = "estimate"
  ) |>
  mutate(
    response = if_else(str_detect(param, "novel"), "novel", "learned"),
    response = factor(response, levels = c("novel", "learned"))
  )
```

### plot drift rate

```{r}
#| label: plot-drift-focused

drift_colors <- c("true" = "salmon", "false" = "steelblue")

# density panel
p_drift_dens <- ggplot(
  drift_plot_data,
  aes(x = estimate, fill = accuracy, colour = accuracy)
) +
  geom_density(alpha = 0.6) +
  scale_fill_manual(values = drift_colors) +
  scale_colour_manual(values = drift_colors) +
  facet_wrap(~condition) +
  labs(
    title = "Distribution of individual drift rates",
    x = "Drift rate", y = "Density",
    fill = "Response type", colour = "Response type"
  ) +
  theme(legend.position = "bottom")

# scatter + smooth — face score
p_drift_face <- ggplot(
  drift_plot_data,
  aes(
    x = face_recog_centred, y = estimate,
    colour = accuracy, fill = accuracy
  )
) +
  geom_point(alpha = 0.3, size = 2) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.4) +
  scale_fill_manual(values = drift_colors) +
  scale_colour_manual(values = drift_colors) +
  facet_wrap(~condition) +
  labs(
    title = "Face recognition effects on drift rate",
    x = "Face recognition ability (centred)",
    y = "Drift rate",
    colour = "Response type", fill = "Response type"
  ) +
  theme(legend.position = "bottom")

# scatter + smooth — car score
p_drift_car <- ggplot(
  drift_plot_data,
  aes(
    x = car_recog_centred, y = estimate,
    colour = accuracy, fill = accuracy
  )
) +
  geom_point(alpha = 0.3, size = 2) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.4) +
  scale_fill_manual(values = drift_colors) +
  scale_colour_manual(values = drift_colors) +
  facet_wrap(~condition) +
  labs(
    title = "Car recognition effects on drift rate",
    x = "Car recognition ability (centred)",
    y = "Drift rate",
    colour = "Response type", fill = "Response type"
  ) +
  theme(legend.position = "bottom")

# combined — replicating image 3 layout
p_drift_combined <- (p_drift_dens | p_drift_face | p_drift_car) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

p_drift_combined



ggsave(here("figures", "corr", "drift_correlations.jpeg"), 
       width = 15, height = 6)

p_drift_face

ggsave(here("figures", "corr", "drift_correlations_simple.jpeg"), 
       width = 8, height = 6)
```

and an alternative

```{r}
#| label: wrangle-drift-alt-style

drift_alt <- theta_pid_q |>
  filter(str_detect(param, "v_S")) |>
  select(pid, param, value) |>
  mutate(
    stimulus = if_else(str_detect(param, "novel"), "novel", "learned"),
    accuracy = if_else(str_detect(param, "TRUE"), "true", "false"),
    condition = factor(
      paste0(stimulus, ".", accuracy),
      levels = c("learned.false", "novel.false", "learned.true", "novel.true")
    ),
    stimulus = factor(stimulus, levels = c("learned", "novel")),
    accuracy = factor(accuracy, levels = c("false", "true"))
  ) |>
  left_join(cam_face |> select(pid, face_recog_centred), by = "pid") |>
  left_join(cam_car  |> select(pid, car_recog_centred),  by = "pid") |>
  rename(face = face_recog_centred, car = car_recog_centred)

head(drift_alt)
```

```{r}
#| label: plot-drift-alt-style
#| fig-width: 10
#| fig-height: 8

# reuse the same plotting functions from the pilot
plot_smooth <- function(data, x, y) {
  ggplot(data, aes(x = {{ x }}, y = {{ y }})) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm") +
    facet_wrap(~condition, nrow = 1) +
    labs(y = "Drift rate (median)")
}

p_face <- plot_smooth(drift_alt, face, value) +
  labs(x = "Face recognition (centred)")
p_car  <- plot_smooth(drift_alt, car,  value) +
  labs(x = "Car recognition (centred)")

p_combi <- p_face / p_car
p_combi

ggsave(here("figures", "corr", "drift_rate_smooth.jpeg"), 
       width = 12, height = 8)
```

and now as a 2x2 structure

```{r}
#| label: plot-drift-2x2
#| fig-width: 10
#| fig-height: 6

plot_2x2 <- function(data, x, y) {
  ggplot(
    data,
    aes(x = {{ x }}, y = {{ y }}, fill = accuracy, colour = accuracy)
  ) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm") +
    scale_fill_brewer(palette = "Dark2") +
    scale_colour_brewer(palette = "Dark2") +
    facet_wrap(~stimulus, nrow = 1) +
    labs(y = "Drift rate (median)", fill = NULL, colour = NULL)
}

## face
p_face_2x2 <- plot_2x2(drift_alt, face, value) +
  labs(x = "Face recognition (centred)", title = "Face") 

## car
p_car_2x2 <- plot_2x2(drift_alt, car, value) +
  labs(x = "Car recognition (centred)", title = "Car")

## combined
(p_face_2x2 | p_car_2x2) +
  plot_layout(guides = "collect",
              axes = "collect") &
  theme(legend.position = "bottom")

ggsave(here("figures", "corr", "drift_rate_2x2.jpeg"), 
       width = 10, height = 7)
```

### plot threshold

```{r}
#| label: plot-threshold-focused

thresh_colors <- c("novel" = "purple", "learned" = "orange")

# density panel
p_thresh_dens <- ggplot(
  thresh_plot_data,
  aes(x = estimate, fill = response, colour = response)
) +
  geom_density(alpha = 0.6) +
  scale_fill_manual(values = thresh_colors) +
  scale_colour_manual(values = thresh_colors) +
  labs(
    title = "Distribution of response thresholds",
    x = "Threshold (B)", y = "Density",
    fill = "Response", colour = "Response"
  ) +
  theme(legend.position = "bottom")

# scatter + smooth — face score
p_thresh_face <- ggplot(
  thresh_plot_data,
  aes(
    x = face_recog_centred, y = estimate,
    colour = response, fill = response
  )
) +
  geom_point(alpha = 0.3, size = 2) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.4) +
  scale_fill_manual(values = thresh_colors) +
  scale_colour_manual(values = thresh_colors) +
  labs(
    title = "Face recognition effects on threshold",
    x = "Face recognition ability (centred)",
    y = "Threshold",
    colour = "Response", fill = "Response"
  ) +
  theme(legend.position = "bottom")

# scatter + smooth — car score
p_thresh_car <- ggplot(
  thresh_plot_data,
  aes(
    x = car_recog_centred, y = estimate,
    colour = response, fill = response
  )
) +
  geom_point(alpha = 0.3, size = 2) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.4) +
  scale_fill_manual(values = thresh_colors) +
  scale_colour_manual(values = thresh_colors) +
  labs(
    title = "Car recognition effects on threshold",
    x = "Car recognition ability (centred)",
    y = "Threshold",
    colour = "Response", fill = "Response"
  ) +
  theme(legend.position = "bottom")

# combined
p_thresh_combined <- (p_thresh_dens | p_thresh_face | p_thresh_car) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

p_thresh_combined

ggsave(here("figures", "corr", "threshold_correlations.jpeg"), 
       width = 15, height = 6)

p_thresh_face
ggsave(here("figures", "corr", "threshold_correlations_simple.jpeg"), 
       width = 8, height = 6)
```

alternative layout

```{r}
#| label: wrangle-threshold-alt-style

thresh_alt <- theta_pid_q |>
  filter(str_detect(param, "B_")) |>
  select(pid, param, value) |>
  mutate(
    response = if_else(str_detect(param, "novel"), "novel", "learned"),
    response = factor(response, levels = c("learned", "novel"))
  ) |>
  left_join(cam_face |> select(pid, face_recog_centred), by = "pid") |>
  left_join(cam_car  |> select(pid, car_recog_centred),  by = "pid") |>
  rename(face = face_recog_centred, car = car_recog_centred)

head(thresh_alt)
```

```{r}
#| label: plot-threshold-smooth
#| fig-width: 10
#| fig-height: 8

plot_smooth_thresh <- function(data, x, y) {
  ggplot(data, aes(x = {{ x }}, y = {{ y }})) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm") +
    facet_wrap(~response, nrow = 1) +
    labs(y = "Threshold (median)")
}

p_thresh_face <- plot_smooth_thresh(thresh_alt, face, value) +
  labs(x = "Face recognition (centred)")
p_thresh_car  <- plot_smooth_thresh(thresh_alt, car,  value) +
  labs(x = "Car recognition (centred)")

p_thresh_combi <- p_thresh_face / p_thresh_car
p_thresh_combi

ggsave(here("figures", "corr", "threshold_smooth.jpeg"), 
       width = 10, height = 7)
```

```{r}
#| label: plot-threshold-2x2
#| fig-width: 8
#| fig-height: 6

plot_2x2_thresh <- function(data, x, y) {
  ggplot(
    data,
    aes(x = {{ x }}, y = {{ y }}, fill = response, colour = response)
  ) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm") +
    scale_fill_manual(values = thresh_colors) +
    scale_colour_manual(values = thresh_colors) +
    facet_wrap(~response, nrow = 1) +
    labs(y = "Threshold (median)", fill = NULL, colour = NULL)
}

p_thresh_face_2x2 <- plot_2x2_thresh(thresh_alt, face, value) +
  labs(x = "Face recognition (centred)", title = "Face")

p_thresh_car_2x2 <- plot_2x2_thresh(thresh_alt, car, value) +
  labs(x = "Car recognition (centred)", title = "Car")

(p_thresh_face_2x2 | p_thresh_car_2x2) +
  plot_layout(guides = "collect", axes = "collect") &
  theme(legend.position = "bottom")

ggsave(here("figures", "corr", "threshold_2x2.jpeg"), 
       width = 10, height = 7)
```

# 5. Conventional correlation table

```{r}
#| label: correlation-table

prereg_params <- c(
  "B_lRnovel", "B_lRlearned", "thresh_diff",
  "v_Snovel_lMTRUE", "v_Snovel_lMFALSE", "discrim_novel",
  "v_Slearned_lMTRUE", "v_Slearned_lMFALSE", "discrim_learned",
  "discrim_diff"
)

param_labels <- c(
  B_lRnovel        = "Threshold: Novel",
  B_lRlearned      = "Threshold: Learned",
  thresh_diff      = "Threshold: Novel - Learned",
  v_Snovel_lMTRUE  = "Drift: Novel (true)",
  v_Snovel_lMFALSE = "Drift: Novel (false)",
  discrim_novel    = "Drift: Novel (true - false)",
  v_Slearned_lMTRUE  = "Drift: Learned (true)",
  v_Slearned_lMFALSE = "Drift: Learned (false)",
  discrim_learned  = "Drift: Learned (true - false)",
  discrim_diff     = "Drift: (Novel - Learned) difference"
)

cor_table <- cor_data |>
  select(all_of(prereg_params), face_recog_centred, car_recog_centred) |>
  pivot_longer(
    cols      = all_of(prereg_params),
    names_to  = "param",
    values_to = "estimate"
  ) |>
  group_by(param) |>
  summarise(
    r_face = cor(estimate, face_recog_centred, use = "complete.obs"),
    r_car  = cor(estimate, car_recog_centred,  use = "complete.obs"),
    .groups = "drop"
  ) |>
  mutate(
    param = factor(param, levels = prereg_params),
    label = param_labels[as.character(param)]
  ) |>
  arrange(param) |>
  select(label, r_face, r_car) |>
  knitr::kable(
    digits    = 3,
    col.names = c("Parameter", "r (Face)", "r (Car)"),
    caption   = "Zero-order correlations between LBA parameters and recognition scores"
  )

cor_table
```

corr mat plot

```{r}
#| label: plot-corrplot
#| fig-width: 12
#| fig-height: 6

cor_matrix_data <- cor_data |>
  select(all_of(prereg_params), face_recog_centred, car_recog_centred) |>
  pivot_longer(
    cols      = all_of(prereg_params),
    names_to  = "param",
    values_to = "estimate"
  ) |>
  group_by(param) |>
  summarise(
    r_face    = cor.test(estimate, face_recog_centred)$estimate,
    lo_face   = cor.test(estimate, face_recog_centred)$conf.int[1],
    hi_face   = cor.test(estimate, face_recog_centred)$conf.int[2],
    r_car     = cor.test(estimate, car_recog_centred)$estimate,
    lo_car    = cor.test(estimate, car_recog_centred)$conf.int[1],
    hi_car    = cor.test(estimate, car_recog_centred)$conf.int[2],
    .groups   = "drop"
  ) |>
  mutate(label = param_labels[param])

# long format for plotting
cor_long_plot <- cor_matrix_data |>
  pivot_longer(
    cols      = c(r_face, r_car, lo_face, hi_face, lo_car, hi_car),
    names_to  = c(".value", "predictor"),
    names_pattern = "(r|lo|hi)_(face|car)"
  ) |>
  mutate(
    predictor  = if_else(predictor == "face", "Face", "Car"),
    predictor  = factor(predictor, levels = c("Face", "Car")),
    label      = factor(label, levels = rev(param_labels)),
    cell_label = sprintf("%.2f\n[%.2f, %.2f]", r, lo, hi)
  )

ggplot(cor_long_plot, aes(x = predictor, y = label, fill = r)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  geom_text(aes(label = cell_label),
            size = 3, fontface = "bold", lineheight = 0.9) +
  scale_fill_distiller(
    palette   = "RdBu",
    limits    = c(-0.3, 0.3),
    direction = 1
  ) +
  labs(
    title = "Zero-order correlations with recognition scores",
    x     = NULL,
    y     = NULL,
    fill  = "r"
  ) +
  theme(
    axis.text.x     = element_text(size = 13, face = "bold"),
    axis.text.y     = element_text(size = 11, face = "bold"),
    legend.position = "right"
  )

ggsave(here("figures", "corr", "corrplot.jpeg"), 
       width = 10, height = 7)
```

# 6. Plausible value correlations

```{r}
#| label: plausible-value-wrangle

all_prereg_params <- c(
  param_cols,
  "discrim_novel", "discrim_learned",
  "discrim_diff", "thresh_diff"
)

# join theta_pid (full posterior) with cambridge scores
# theta_pid has 300 pids x 9 params x 15000 draws
pv_data <- theta_pid |>
  pivot_wider(names_from = param, values_from = value) |>
  mutate(
    discrim_novel   = v_Snovel_lMTRUE - v_Snovel_lMFALSE,
    discrim_learned = v_Slearned_lMTRUE - v_Slearned_lMFALSE,
    discrim_diff    = discrim_novel - discrim_learned,
    thresh_diff     = B_lRnovel - B_lRlearned
  ) |>
  left_join(
    cam_face |> select(pid, face_recog_centred),
    by = "pid"
  ) |>
  left_join(
    cam_car |> select(pid, car_recog_centred),
    by = "pid"
  ) |>
  pivot_longer(
    cols      = all_of(all_prereg_params),
    names_to  = "param",
    values_to = "estimate"
  )

# compute correlation for each draw x param x predictor
# this will take a moment with 15000 draws
pv_cors <- pv_data |>
  group_by(.draw, param) |>
  summarise(
    r_face = cor(estimate, face_recog_centred, use = "complete.obs"),
    r_car = cor(estimate, car_recog_centred, use = "complete.obs"),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols      = c(r_face, r_car),
    names_to  = "predictor",
    values_to = "r"
  ) |>
  mutate(
    predictor = if_else(predictor == "r_face", "Face", "Car"),
    predictor = factor(predictor, levels = c("Face", "Car"))
  )

# save out — this is a big object
write_csv(pv_cors, here("data", "pv_cors.csv"))
```

```{r}
#| label: plausible-value-summary

# summarise
pv_summary <- pv_cors |>
  group_by(param, predictor) |>
  median_qi(r, .width = c(0.95)) |>
  mutate(label = param_labels[param]) |> 
  print(n=Inf)
```

```{r}
#| label: plot-pv-key-params
#| fig-width: 12
#| fig-height: 8

# key params per prereg: drift discriminations + threshold + their differences
key_params <- c(
  "discrim_novel", "discrim_learned", "discrim_diff",
  "B_lRnovel", "B_lRlearned", "thresh_diff"
)

key_labels <- c(
  discrim_novel   = "Drift: Novel (true-false)",
  discrim_learned = "Drift: Learned (true-false)",
  discrim_diff    = "Drift: Novel - Learned",
  B_lRnovel       = "Threshold: Novel",
  B_lRlearned     = "Threshold: Learned",
  thresh_diff     = "Threshold: Novel - Learned"
)

pv_cors_key <- pv_cors |>
  filter(param %in% key_params) |>
  mutate(
    param = factor(param, levels = key_params,
                   labels = key_labels)
  )

ggplot(pv_cors_key, aes(x = r, y = predictor, 
                         fill = predictor, colour = predictor)) +
  stat_halfeye(alpha = 0.7, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  scale_colour_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  facet_wrap(~param, ncol = 3) +
  labs(
    title = "Posterior correlation distributions — key parameters",
    x     = "Plausible correlation with recognition score",
    y     = NULL,
    fill  = NULL, colour = NULL
  ) +
  theme(legend.position = "bottom")

ggsave(here("figures", "pv_cors_key.jpeg"), width = 12, height = 8)
```

```{r}
#| label: plot-pv-threshold

thresh_colors <- c("Novel" = safe_pal[1], "Learned" = safe_pal[2])

# left panel — novel and learned
pv_thresh_main <- pv_cors |>
  filter(param %in% c("B_lRnovel", "B_lRlearned")) |>
  mutate(
    response  = if_else(param == "B_lRnovel", "Novel", "Learned"),
    response  = factor(response, levels = c("Novel", "Learned"))
  )

p_thresh_left <- ggplot(
  pv_thresh_main,
  aes(x = r, y = predictor, fill = response, colour = response)
) +
  stat_halfeye(alpha = 0.7, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = thresh_colors) +
  scale_colour_manual(values = thresh_colors) +
  labs(
    x = "Plausible r", y = NULL,
    title = "Threshold",
    fill = NULL, colour = NULL
  ) +
  theme(legend.position = "bottom")

# right panel — difference
pv_thresh_diff <- pv_cors |>
  filter(param == "thresh_diff")

p_thresh_right <- ggplot(
  pv_thresh_diff,
  aes(x = r, y = predictor)
) +
  stat_halfeye(
    alpha = 0.7, .width = c(0.5, 0.95),
    fill = "#1A7A7A"
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  labs(
    x = "Plausible r", y = NULL,
    title = "Threshold (Novel - Learned)",
    fill = NULL, colour = NULL
  ) +
  theme(legend.position = "bottom")

p_thresh <- (p_thresh_left | p_thresh_right) +
  plot_layout(guides = "collect",
              axes = "collect") &
  theme(legend.position = "bottom")

p_thresh

ggsave(here("figures", "corr", "pv_threshold.jpeg"), 
       width = 10, height = 5)
```

```{r}
#| label: plot-pv-drift

drift_colors <- c("Novel" = safe_pal[1], "Learned" = safe_pal[2])

# left panel — true and false per condition
pv_drift_main <- pv_cors |>
  filter(param %in% c(
    "v_Snovel_lMTRUE", "v_Snovel_lMFALSE",
    "v_Slearned_lMTRUE", "v_Slearned_lMFALSE"
  )) |>
  mutate(
    condition = if_else(str_detect(param, "novel"), "Novel", "Learned"),
    accuracy  = if_else(str_detect(param, "TRUE"), "true", "false"),
    condition = factor(condition, levels = c("Novel", "Learned")),
    accuracy  = factor(accuracy, levels = c("true", "false"))
  )

p_drift_left <- ggplot(
  pv_drift_main,
  aes(
    x = r, y = predictor, fill = condition, colour = condition,
    alpha = accuracy
  )
) +
  stat_halfeye(.width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = drift_colors) +
  scale_colour_manual(values = drift_colors) +
  scale_alpha_manual(values = c("true" = 0.8, "false" = 0.3)) +
  labs(
    x = "Plausible r", y = NULL,
    title = "Drift rate (true / false)",
    fill = NULL, colour = NULL, alpha = "Accuracy"
  ) +
  theme(legend.position = "bottom")

# middle panel — discrimination (true - false) per condition
pv_drift_discrim <- pv_cors |>
  filter(param %in% c("discrim_novel", "discrim_learned")) |>
  mutate(
    condition = if_else(param == "discrim_novel", "Novel", "Learned"),
    condition = factor(condition, levels = c("Novel", "Learned"))
  )

p_drift_mid <- ggplot(
  pv_drift_discrim,
  aes(x = r, y = predictor, fill = condition, colour = condition)
) +
  stat_halfeye(alpha = 0.7, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = drift_colors) +
  scale_colour_manual(values = drift_colors) +
  labs(
    x = "Plausible r", y = NULL,
    title = "Drift (true - false)",
    fill = NULL, colour = NULL
  ) +
  theme(legend.position = "bottom")

# right panel — novel-learned difference of discriminations
pv_drift_diff <- pv_cors |>
  filter(param == "discrim_diff")

p_drift_right <- ggplot(
  pv_drift_diff,
  aes(x = r, y = predictor, fill = predictor, colour = predictor)
) +
  stat_halfeye(alpha = 0.7, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  scale_colour_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  labs(
    x = "Plausible r", y = NULL,
    title = "Drift difference\n(Novel - Learned)",
    fill = NULL, colour = NULL
  ) +
  theme(legend.position = "bottom")

p_drift <- (p_drift_left | p_drift_mid | p_drift_right) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

p_drift

ggsave(here("figures", "pv_drift.jpeg"), 
       width = 14, height = 5)
```

```{r}
#| label: plot-pv-drift-v2
#| fig-width: 14
#| fig-height: 6

drift_colors <- c("Novel" = safe_pal[1], "Learned" = safe_pal[2])

# left panel — true and false per condition, faceted by predictor
pv_drift_main <- pv_cors |>
  filter(param %in% c(
    "v_Snovel_lMTRUE", "v_Snovel_lMFALSE",
    "v_Slearned_lMTRUE", "v_Slearned_lMFALSE"
  )) |>
  mutate(
    condition = if_else(str_detect(param, "novel"), "Novel", "Learned"),
    accuracy  = if_else(str_detect(param, "TRUE"), "true", "false"),
    condition = factor(condition, levels = c("Learned", "Novel")),
    accuracy  = factor(accuracy, levels = c("true", "false"))
  )

p_drift_left <- ggplot(
  pv_drift_main,
  aes(
    x = r, y = condition, fill = condition, colour = condition,
    alpha = accuracy
  )
) +
  stat_halfeye(.width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = drift_colors) +
  scale_colour_manual(values = drift_colors, guide = "none") +
  scale_alpha_manual(values = c("true" = 0.8, "false" = 0.3), guide = "none") +
  facet_wrap(~predictor, ncol = 1) +
  labs(
    x = "Plausible r", y = NULL,
    title = "Drift rate",
    subtitle = "light = false; dark = true",
    fill = NULL
  ) +
  theme(legend.position = "none")

# middle panel — discrimination (true - false) per condition, faceted by predictor
pv_drift_discrim <- pv_cors |>
  filter(param %in% c("discrim_novel", "discrim_learned")) |>
  mutate(
    condition = if_else(param == "discrim_novel", "Novel", "Learned"),
    condition = factor(condition, levels = c("Learned", "Novel"))
  )

p_drift_mid <- ggplot(
  pv_drift_discrim,
  aes(x = r, y = condition, fill = condition, colour = condition)
) +
  stat_halfeye(alpha = 0.7, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = drift_colors) +
  scale_colour_manual(values = drift_colors) +
  facet_wrap(~predictor, ncol = 1) +
  labs(
    x = "Plausible r", y = NULL,
    title = "Drift (true - false)",
    fill = NULL, colour = NULL
  ) +
  theme(legend.position = "bottom")

# right panel — novel-learned difference, faceted by predictor
pv_drift_diff <- pv_cors |>
  filter(param == "discrim_diff") |>
  mutate(
    condition = "Novel - Learned",
    condition = factor(condition)
  )

p_drift_right <- ggplot(
  pv_drift_diff,
  aes(x = r, y = "")
) +
  stat_halfeye(alpha = 0.7, .width = c(0.5, 0.95),
               fill = "#1A7A7A") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  facet_wrap(~predictor, ncol = 1) +
  labs(
    x = "Plausible r", y = NULL,
    title = "Drift difference\n(Novel - Learned)",
    fill = NULL, colour = NULL
  ) +
  theme(legend.position = "none")

# combine
p_drift <- (p_drift_left | p_drift_mid | p_drift_right) +
  plot_layout(axes = "collect") 

p_drift

ggsave(here("figures", "corr", "pv_drift.jpeg"), 
       width = 12, height = 6)
```

