---
title: "wrangle"
author: "Raph & Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

This file wrangles data for the individual differences face familiarity study.

It produces two cleaned datasets:

- `data/data.csv` — face recognition task.
- `data/data_cambridge_scores.csv` — Cambridge face/car memory tests.

Exclusion decisions follow the pre-registered criteria.

It also produces one combined raw dataset:

- `data/raw_data.csv` — face recognition task.

```{r}
#| label: setup
#| include: false

pkgs <- c(
  "tidyverse", "here", "patchwork", "paletteer", "ggpubr"
  )
vapply(
  pkgs, library, logical(1), character.only = TRUE, logical.return = TRUE
  )

theme_set(
  theme_bw() +
    theme(
      text  = element_text(size = 15, face = "bold"),
      title = element_text(size = 15, face = "bold"),
      legend.position = "none"
    )
)

safe_pal  <- paletteer::paletteer_d("rcartocolor::Safe", 4)

mycolors <- c("novel" = safe_pal[1], "learned" = safe_pal[2])

pd2 <- position_dodge(1)
```

# 1. Read in data

## read in the raw data

Combine individual CSV files (run once; output cached to `data/raw_data.csv`).

```{r}
#| label: read-raw
#| eval: false   # set TRUE on first run to regenerate raw_data.csv

raw <- list.files(
  path       = here("data", "raw"),
  pattern    = "^.*\\.csv$",
  full.names = TRUE,
  recursive  = TRUE
) |>
  set_names() |>
  map(read_csv, show_col_types = FALSE) |>
  list_rbind(names_to = "file")

write_csv(raw, here("data", "raw_data.csv"))
```

```{r}
#| label: read-combined

raw <- read_csv(here("data", "raw_data.csv"), 
                show_col_types = FALSE)
head(raw)
glimpse(raw)
```

## read in the clean data

If you read in the clean data below, then skip to section 6 and 7 for summary and plotting.

```{r}
#| label: read-clean
#| eval: false

## main task RT and acc data
data_clean <- read_csv("data/data.csv")
head(data_clean)
glimpse(data_clean)

## Cambridge face and car memory task data
cambridge_tests <- read_csv("data/cambridge_scores.csv")
head(cambridge_tests)
glimpse(cambridge_tests)

## Add the gender for clean Cambridge scores
demo <- read_csv("data/demo.csv")
cambridge_tests <- cambridge_tests |>
  left_join(demo, Sex, by = "pid") |>
  mutate(Sex = as.factor(Sex))

## Save Cambridge scores

write_csv(cambridge_tests, here("data/cambridge_scores.csv"))
```

---

# 2. Wrangle: face recognition task

```{r}
#| label: wrangle-face

# NOTE: block assignment via rep() assumes rows are already in trial order
# within each pid. Verify this holds after read_csv / list_rbind.
n_blocks      <- 3
trials_per_block <- 60   # 180 trials total per participant

data <- raw |>
  filter(!is.na(recognition_loop.thisRepN)) |>
  rename(
    keypress = response_recog.keys,
    acc      = response_recog.corr,
    rt       = response_recog.rt
  ) |>
  select(participant, keypress, acc, rt) |>
  mutate(
    pid      = sprintf("%03d", as.integer(factor(participant))),
    keypress = as.factor(keypress),
    block    = rep(
      seq_len(n_blocks),
      each  = trials_per_block,
      times = length(unique(pid))
    ),
    # cond is fully determined by acc + key mapping:
    # 'q' = Learned response; 'p' = Novel response
    # correct Learned trial OR incorrect Novel trial → Learned stimulus
    cond = if_else(
      (keypress == "q" & acc == 1) | (keypress == "p" & acc == 0),
      "learned", "novel"
    ),
    cond = factor(cond, levels = c("learned", "novel"))
  ) |>
  relocate(pid, block, cond, .before = keypress) |>
  select(-participant) |> 
  arrange(pid)

head(data)
glimpse(data)

cat("Participants before exclusions:", length(unique(data$pid)), "\n")
```

---

# 3. Exclusions: face recognition task

Each stage below tracks how many participants are removed and why.

```{r}
#| label: exclusion-tracking

# Helper: count remaining pids
n_pids <- function(df) length(unique(df$pid))

# Running exclusion log — we'll append to this
excl_log <- tibble(
  stage     = character(),
  criterion = character(),
  n_removed = integer(),
  n_remain  = integer()
)

log_exclusion <- function(log, stage, criterion, before, after) {
  bind_rows(log, tibble(
    stage     = stage,
    criterion = criterion,
    n_removed = as.integer(before - after),
    n_remain  = as.integer(after)
  ))
}

n_start <- n_pids(data)
```

## 3.1 Missed trials (> 40% NA rt)

```{r}
#| label: excl-missed

total_trials   <- n_blocks * trials_per_block  # 180
missed_thresh  <- 0.40

missed_summary <- data |>
  group_by(pid) |>
  summarise(n_missed = sum(is.na(rt)), .groups = "drop") |>
  mutate(prop_missed = n_missed / total_trials)

pids_missed <- missed_summary |>
  filter(prop_missed > missed_thresh) |>
  pull(pid)

data_excl1 <- data |> filter(!pid %in% pids_missed)

excl_log <- log_exclusion(
  excl_log, "1", 
  glue::glue("> {missed_thresh * 100}% missed trials"),
  n_start, n_pids(data_excl1)
)

# Visualise missed-trial distribution
missed_trials_plot <- missed_summary |>
  mutate(excluded = pid %in% pids_missed) |>
  ggplot(aes(x = prop_missed, fill = excluded)) +
  geom_histogram(binwidth = 0.02, colour = "white") +
  geom_rug(data = ~filter(., excluded), 
           aes(colour = excluded), sides = "b", linewidth = 1) +
  geom_vline(xintercept = missed_thresh, linetype = "dashed", colour = "red") +
  scale_fill_manual(values = c("FALSE" = "steelblue", "TRUE" = "tomato"),
                    labels = c("Retained", "Excluded")) +
  labs(
    title = "Proportion of missed trials",
    x     = "Proportion missed",
    y     = "Count",
    fill  = NULL
  )
missed_trials_plot
```

## 3.2 RT range violations (> 25% of trials outside 0.2–2.0 s)

```{r}
#| label: excl-rt-range

rt_lo   <- 0.2
rt_hi   <- 2.0
rt_thresh <- 0.25   # > 25% violations → exclude

rt_check <- data_excl1 |>
  filter(!is.na(rt)) |>
  group_by(pid) |>
  summarise(
    n_trials      = n(),
    n_out_of_range = sum(rt < rt_lo | rt > rt_hi),
    prop_out       = n_out_of_range / n_trials,
    .groups        = "drop"
  )

pids_rt <- rt_check |>
  filter(prop_out > rt_thresh) |>
  pull(pid)

data_excl2 <- data_excl1 |> filter(!pid %in% pids_rt)

excl_log <- log_exclusion(
  excl_log, "2",
  glue::glue("> {rt_thresh * 100}% RT outside [{rt_lo}, {rt_hi}] s"),
  n_pids(data_excl1), n_pids(data_excl2)
)

# Visualise RT violation distribution
rt_check_plot <- rt_check |>
  mutate(excluded = pid %in% pids_rt) |>
  ggplot(aes(x = prop_out, fill = excluded)) +
  geom_histogram(binwidth = 0.02, colour = "white") +
  geom_rug(data = ~filter(., excluded), 
           aes(colour = excluded), sides = "b", linewidth = 1) +
  geom_vline(xintercept = rt_thresh, linetype = "dashed", colour = "red") +
  scale_fill_manual(values = c("FALSE" = "steelblue", "TRUE" = "tomato"),
                    labels = c("Retained", "Excluded")) +
  labs(
    title = "Proportion of RTs",
    x     = "Proportion out of range",
    y     = "Count",
    fill  = NULL
  )
rt_check_plot
```

## 3.3 Accuracy floor (< 55%) and ceiling (> 95%)

```{r}
#| label: excl-accuracy

acc_lo <- 55
acc_hi <- 95

# Remove out-of-range RTs before accuracy check (trial-level filter)
data_valid_rt <- data_excl2 |>
  filter(!is.na(rt), rt >= rt_lo, rt <= rt_hi)

acc_check <- data_valid_rt |>
  group_by(pid) |>
  summarise(
    n_trials = n(),
    n_correct = sum(acc),
    perc_acc  = n_correct / n_trials * 100,
    .groups   = "drop"
  )

pids_acc_floor   <- acc_check |> filter(perc_acc < acc_lo) |> pull(pid)
pids_acc_ceiling <- acc_check |> filter(perc_acc > acc_hi) |> pull(pid)
pids_acc         <- union(pids_acc_floor, pids_acc_ceiling)

data_excl3 <- data_valid_rt |> filter(!pid %in% pids_acc)

excl_log <- log_exclusion(
  excl_log, "3a",
  glue::glue("Accuracy < {acc_lo}%"),
  n_pids(data_valid_rt), n_pids(data_valid_rt) - length(pids_acc_floor)
)
excl_log <- log_exclusion(
  excl_log, "3b",
  glue::glue("Accuracy > {acc_hi}%"),
  n_pids(data_valid_rt) - length(pids_acc_floor),
  n_pids(data_excl3)
)

# Visualise accuracy distribution
acc_check_plot <- acc_check |>
  mutate(
    excluded = case_when(
      perc_acc < acc_lo ~ "Floor",
      perc_acc > acc_hi ~ "Ceiling",
      TRUE              ~ "Retained"
    ),
    excluded = factor(excluded, levels = c("Retained", "Floor", "Ceiling"))
  ) |>
  ggplot(aes(x = perc_acc, fill = excluded)) +
  geom_histogram(binwidth = 2, colour = "white") +
  geom_rug(data = ~filter(., excluded %in% c("Floor", "Ceiling")), 
           aes(colour = excluded), sides = "b", linewidth = 0.25) +
  geom_vline(xintercept = c(acc_lo, acc_hi), linetype = "dashed", colour = "red") +
  scale_fill_manual(
    values = c("Retained" = "steelblue", "Floor" = "tomato", "Ceiling" = "orange")
  ) +
  labs(
    title = "Accuracy",
    x     = "Accuracy (%)",
    y     = "Count",
    fill  = NULL
  )
acc_check_plot
```

## 3.4 Exclusion summary

```{r}
#| label: excl-summary

excl_log |>
  knitr::kable(
    col.names = c("Stage", "Criterion", "N removed", "N remaining"),
    caption   = "Participant exclusions — face recognition task"
  )

cat(
  "\nFinal N (face recognition task):", n_pids(data_excl3), "\n",
  "Started with:", n_start, "\n",
  "Total excluded:", n_start - n_pids(data_excl3), "\n"
)

## and plot
p_exclusions <- missed_trials_plot | rt_check_plot | acc_check_plot
p_exclusions

ggsave("figures/descriptive/data_exclusions.png",
       width = 15, height = 7)
```

```{r}
#| label: finalise-face-data

data_clean <- data_excl3
  
#write_csv(data_clean, "data/data.csv")
```

---

# 4. Wrangle: Cambridge memory tests

```{r}
#| label: wrangle-cambridge

order_exp <- raw |>
  filter(!is.na(Title)) |>
  select(participant, Title) |>
  mutate(pid   = sprintf("%03d", as.integer(factor(participant))),
         Title = as.factor(Title)) |>
  select(-participant) |> 
  group_by(pid) |>
  mutate(trial_block = row_number()) |>
  ungroup() |>
  mutate(
    task = fct_recode(factor(Title),
                      face = "Cambridge Face Memory Test",
                      car  = "Cambridge Car Memory Test")
  ) |>
  select(pid, trial_block, task)

cambridge_tests <- raw |>
  filter(!is.na(key_resp_12.corr)) |>
  rename(
    keypress = key_resp_12.keys,
    acc      = key_resp_12.corr,
    rt       = key_resp_12.rt
  ) |>
  mutate(pid = sprintf("%03d", as.integer(factor(participant)))) |>
  group_by(pid) |>
  mutate(
    trial_block = if_else(row_number() <= 54, 1L, 2L),
    stim        = factor(if_else((row_number() - 1) %% 54 < 30, "no_noise", "noise"))
  ) |>
  ungroup() |>
  left_join(order_exp, by = c("pid", "trial_block")) |>
  select(pid, task, stim, keypress, acc, rt) |>
  filter(
    !pid %in% pids_missed,
    !pid %in% pids_rt,
    !pid %in% pids_acc
  ) |>
  arrange(pid)

cat("Participants in Cambridge data:", n_pids(cambridge_tests), "\n")
```

some checks

```{r}
#| label: cambridge-checks

# 1. Each pid should have exactly 108 trials (before exclusions)
cambridge_tests |> count(pid) |> count(n)

# 2. Each pid x task should have exactly 54 trials per task
cambridge_tests |> count(pid, task) |> count(task, n)

# 3. stim split should be 30 no_noise + 24 noise per pid x task
cambridge_tests |> count(pid, task, stim) |> count(task, stim, n)
```

---

# 5. Save cleaned data

```{r}
#| label: save-data

write_csv(data_clean, here("data", "data.csv"))
write_csv(cambridge_tests, here("data", "cambridge_scores.csv"))
```

---

# 6. Summary statistics

Helper for computing mean ± SEM at any grouping level:

```{r}
#| label: summary-helpers

summarise_rt <- function(df, ...) {
  df |>
    drop_na(rt) |>
    group_by(...) |>
    summarise(
      mean = mean(rt),
      sd   = sd(rt),
      n    = n(),
      .groups = "drop"
    ) |>
    mutate(
      sem = sd / sqrt(n),
      ci  = sem * 1.96
    )
}

summarise_acc <- function(df, ...) {
  df |>
    group_by(...) |>
    summarise(
      n_correct = sum(acc),
      n         = n(),
      .groups   = "drop"
    ) |>
    mutate(perc = n_correct / n * 100)
}

aggregate_acc <- function(acc_pid_df, ...) {
  acc_pid_df |>
    group_by(...) |>
    summarise(
      perc_mean = mean(perc),
      perc_sd   = sd(perc),
      n         = n_distinct(pid),
      .groups   = "drop"
    ) |>
    mutate(
      perc = perc_mean,
      sd   = perc_sd,
      sem  = sd / sqrt(n)
    ) |>
    select(-perc_mean, -perc_sd)
}
```

## Face recognition task

```{r}
#| label: summaries-face

# RT
rt_pid   <- summarise_rt(data_clean, pid, block, cond)
rt_group <- summarise_rt(data_clean, block, cond)

# Accuracy
acc_pid   <- summarise_acc(data_clean, pid, block, cond)
acc_group <- aggregate_acc(acc_pid, block, cond)

# Collapsed across blocks
rt_pid_cond   <- summarise_rt(data_clean, pid, cond)
rt_group_cond <- summarise_rt(data_clean, cond)

acc_pid_cond   <- summarise_acc(data_clean, pid, cond)
acc_group_cond <- aggregate_acc(acc_pid_cond, cond)
```

## Cambridge tests

```{r}
#| label: summaries-cambridge

rt_pid_camb   <- summarise_rt(cambridge_tests, pid, task, stim)
rt_group_camb <- summarise_rt(cambridge_tests, task, stim)

acc_pid_camb   <- summarise_acc(cambridge_tests, pid, task, stim, Sex)
acc_group_camb <- aggregate_acc(acc_pid_camb, task, stim)
acc_group_camb_gender <- aggregate_acc(acc_pid_camb, task, Sex, stim)

# Collapsed across stim
rt_pid_camb_task   <- summarise_rt(cambridge_tests, pid, task)
rt_group_camb_task <- summarise_rt(cambridge_tests, task)

acc_pid_camb_task   <- summarise_acc(cambridge_tests, pid, task, Sex)
acc_group_camb_task <- aggregate_acc(acc_pid_camb_task, task)
acc_group_camb_task_gender <-aggregate_acc(acc_pid_camb, task, Sex)

```

## save out summary data

```{r}
#| label: save-summaries-cambridge

write_csv(acc_pid_camb_task, "data/cambridge_acc_scores_pid.csv")
```


---

# 7. Plots

## 7.1 Face recognition — RT and accuracy distributions

```{r}
#| label: plot-face-density
#| fig-width: 10
#| fig-height: 5

p_rt_dens <- ggplot(data_clean, aes(x = rt, fill = cond, colour = cond)) +
  geom_density(alpha = 0.4) +
  scale_fill_manual(values = mycolors) +
  scale_colour_manual(values = mycolors) +
  facet_wrap(~block, labeller = label_both) +
  xlim(0, 3) +
  labs(title = "RT distribution by block and condition",
       x = "Response time (s)", y = "Density", fill = NULL, colour = NULL)

p_acc_bar <- data_clean |>
  group_by(block, cond) |>
  summarise(perc = mean(acc) * 100, .groups = "drop") |>
  ggplot(aes(x = cond, y = perc, fill = cond, colour = cond)) +
  geom_col(alpha = 0.7) +
  scale_fill_manual(values = mycolors) +
  scale_colour_manual(values = mycolors) +
  facet_wrap(~block, labeller = label_both) +
  ylim(0, 100) +
  labs(title = "Accuracy by block and condition",
       x = NULL, y = "Accuracy (%)", fill = NULL, colour = NULL)

p_rt_dens / p_acc_bar

ggsave("figures/descriptive/rt_acc_block.png",
       width = 8, height = 6)
```

## 7.2 Face recognition — group-level RT and accuracy by block

```{r}
#| label: plot-face-group-block
#| fig-width: 10
#| fig-height: 5

p_rt_block <- ggplot(rt_group, aes(x = block, y = mean, colour = cond, fill = cond)) +
  geom_line(aes(group = cond), position = pd2) +
  geom_point(position = pd2, size = 2) +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem),
                width = 0.2, position = pd2) +
  scale_fill_manual(values = mycolors) +
  scale_colour_manual(values = mycolors) +
  ylim(0, 2) +
  labs(title = "Mean RT by block", x = "Block",
       y = "Mean RT (s)", colour = NULL, fill = NULL)

p_acc_block <- ggplot(acc_group, aes(x = block, y = perc, colour = cond, fill = cond)) +
  geom_line(aes(group = cond), position = pd2) +
  geom_point(position = pd2, size = 2) +
  geom_errorbar(aes(ymin = perc - sem, ymax = perc + sem),
                width = 0.2, position = pd2) +
  scale_fill_manual(values = mycolors) +
  scale_colour_manual(values = mycolors) +
  ylim(0, 100) +
  labs(title = "Mean accuracy by block", x = "Block",
       y = "Accuracy (%)", colour = NULL, fill = NULL)

p_rt_block | p_acc_block

ggsave("figures/descriptive/rt_acc_block_means.png",
       width = 8, height = 6)
```

## 7.3 Face recognition — collapsed across blocks (pid dots + group mean)

```{r}
#| label: plot-face-group-cond
#| fig-width: 8
#| fig-height: 5

p_rt_cond <- ggplot(rt_group_cond, aes(x = cond, y = mean, fill = cond, colour = cond)) +
  geom_col(alpha = 0.7, position = pd2) +
  geom_point(data = rt_pid_cond, aes(y = mean),
             alpha = 0.4, position = position_jitter(width = 0.1)) +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem),
                colour = "black", width = 0.2, position = pd2) +
  scale_fill_manual(values = mycolors) +
  scale_colour_manual(values = mycolors) +
  ylim(0, 2) +
  labs(title = "RT by condition", x = NULL, y = "Mean RT (s)")

p_acc_cond <- ggplot(acc_group_cond, aes(x = cond, y = perc, fill = cond, colour = cond)) +
  geom_col(alpha = 0.7, position = pd2) +
  geom_point(data = acc_pid_cond, aes(y = perc),
             alpha = 0.4, position = position_jitter(width = 0.1)) +
  geom_errorbar(aes(ymin = perc - sem, ymax = perc + sem),
                colour = "black", width = 0.2, position = pd2) +
  scale_fill_manual(values = mycolors) +
  scale_colour_manual(values = mycolors) +
  ylim(0, 100) +
  labs(title = "Accuracy by condition", x = NULL, y = "Accuracy (%)")

p_rt_cond | p_acc_cond

ggsave("figures/descriptive/rt_acc_bar_dots.png",
       width = 8, height = 6)
```

## 7.4 Face recognition — acc and rt paired plots

```{r}
#| label: plot-face-paired
#| fig-width: 8
#| fig-height: 5

## paired plots for summarised RT and accuracy data
paired_plot <- function(df, yvar, ylab, ylims) {
  
  # set consistent jitter for dots and lines
  pos <- position_jitter(width = 0.06, height = 0, seed = 123)
  
  ggplot(df, aes(x = cond, y = {{ yvar }}, 
                 fill = cond, colour = cond)) +
    # per-participant lines
    geom_line(
      aes(group = pid),
      position = pos,
      colour    = "grey70",
      linewidth = 0.3,
      alpha     = 0.5
    ) +
    # per-participant dots
    geom_point(
      position = pos,
      alpha    = 0.5,
      size     = 1.5
    ) +
    # group mean + 95% CI
    stat_summary(
      fun.data  = mean_se,
      geom      = "pointrange",
      colour    = "black",
      size      = 0.4,
      linewidth = 1.2,
      fun.args  = list(mult = 1.96)
    ) +
    # line connecting group means
    stat_summary(
      aes(group = 1),
      fun       = mean,
      geom      = "line",
      colour    = "black",
      linewidth = 0.8
    ) +
    scale_fill_manual(values   = mycolors) +
    scale_colour_manual(values = mycolors) +
    coord_cartesian(ylim = ylims) +
    labs(x = NULL, y = ylab) 
}

acc_paired <- paired_plot(acc_pid_cond, perc, "Accuracy [%]",  c(0, 100))

rt_paired  <- paired_plot(rt_pid_cond,  mean, "Mean RT [s]",   c(0, 1.75))

acc_paired | rt_paired

ggsave("figures/descriptive/rt_acc_paired_plots.png",
       width = 8, height = 6)
```

## 7.5 Cambridge tests — RT and accuracy by task and stim type

```{r}
#| label: plot-cambridge-stim
#| fig-width: 10
#| fig-height: 5

p_rt_camb <- ggplot(rt_group_camb, aes(x = stim, y = mean, fill = stim, colour = stim)) +
  geom_col(alpha = 0.7, position = pd2) +
  geom_point(data = rt_pid_camb, aes(y = mean),
             alpha = 0.4, position = position_jitter(width = 0.1, height=0)) +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem),
                colour = "black", width = 0.2, position = pd2) +
  facet_wrap(~task) +
  ylim(0, 20) +
  labs(title = "Cambridge RT by task and stimulus", x = NULL,
       y = "Mean RT (s)", fill = NULL, colour = NULL)

p_acc_camb <- ggplot(acc_group_camb, aes(x = stim, y = perc, fill = stim, colour = stim)) +
  geom_col(alpha = 0.7, position = pd2) +
  geom_point(data = acc_pid_camb, aes(y = perc),
             alpha = 0.4, position = position_jitter(width = 0.1, height=0)) +
  geom_errorbar(aes(ymin = perc - sem, ymax = perc + sem),
                colour = "black", width = 0.2, position = pd2) +
  facet_wrap(~task) +
  ylim(0, 100) +
  labs(title = "Cambridge accuracy by task and stimulus", x = NULL,
       y = "Accuracy (%)", fill = NULL, colour = NULL)

p_rt_camb | p_acc_camb

ggsave("figures/descriptive/camb_tests_by_noise.png",
       width = 10, height = 7)
```

```{r}
#| label: plot-cambridge-stim-density
#| fig-width: 10
#| fig-height: 5

p_rt_camb_dens <- ggplot(rt_pid_camb, aes(x = mean, fill = stim, colour = stim)) +
  geom_density(alpha = 0.7) +
  geom_rug() +
  facet_wrap(~task) +
  labs(title = "Cambridge RT by task and stimulus",
       x = "Mean RT (s)", fill = NULL, colour = NULL) +
  theme(legend.position = "bottom")

p_acc_camb_dens <- ggplot(acc_pid_camb, aes(x = perc, fill = stim, colour = stim)) +
  geom_density(alpha = 0.7) +
  geom_rug() +
  facet_wrap(~task) +
  labs(title = "Cambridge accuracy by task and stimulus",
       x = "Accuracy (%)", fill = NULL, colour = NULL) +
  theme(legend.position = "bottom")

(p_rt_camb_dens | p_acc_camb_dens) +
  plot_layout(axes = "collect",
              guides = "collect") &
  theme(legend.position = "bottom")

ggsave("figures/descriptive/camb_tests_by_noise_density.png",
       width = 10, height = 7)
```

## Cambridge tests — accuracy by task and gender
```{r}
p_acc_camb_gender <- ggplot(acc_group_camb_task_gender, aes(x = Sex, y = perc, fill = Sex, colour = Sex)) +
  geom_col(alpha = 0.7, position = pd2) +
  geom_point(data = acc_pid_camb, aes(y = perc),
             alpha = 0.4, position = position_jitter(width = 0.1, height=0)) +
  geom_errorbar(aes(ymin = perc - sem, ymax = perc + sem),
                colour = "black", width = 0.2, position = pd2) +
  stat_compare_means(data = acc_pid_camb,
                     aes(x = Sex, y = perc),
                     method = "t.test",
                     comparisons = list(c("Female", "Male")),
                     label = "p.signif",   # or "p.format" for exact p-value
                     colour = "black") +
  facet_wrap(~task) +
  ylim(0, 110) +
  scale_y_continuous(breaks=c(0,20,40,60,80,100)) +
  labs(title = "Cambridge accuracy by task and gender", x = NULL,
       y = "Accuracy (%)", fill = NULL, colour = NULL)


ggsave("figures/descriptive/camb_tests_acc_by_gender.png",
       width = 10, height = 7)
```


## 7.6 Cambridge tests — collapsed across stim

```{r}
#| label: plot-cambridge-task
#| fig-width: 8
#| fig-height: 5

p_rt_camb_task <- ggplot(rt_group_camb_task,
                          aes(x = task, y = mean, fill = task, colour = task)) +
  geom_col(alpha = 0.7) +
  geom_point(data = rt_pid_camb_task, aes(y = mean),
             alpha = 0.4, position = position_jitter(width = 0.1, height=0)) +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem),
                colour = "black", width = 0.2) +
  ylim(0, 20) +
  labs(title = "Cambridge RT by task", x = NULL, y = "Mean RT (s)")

p_acc_camb_task <- ggplot(acc_group_camb_task,
                           aes(x = task, y = perc, fill = task, colour = task)) +
  geom_col(alpha = 0.7) +
  geom_point(data = acc_pid_camb, aes(y = perc),
             alpha = 0.4, position = position_jitter(width = 0.1, height=0)) +
  geom_errorbar(aes(ymin = perc - sem, ymax = perc + sem),
                colour = "black", width = 0.2) +
  ylim(0, 100) +
  labs(title = "Cambridge accuracy by task", x = NULL, y = "Accuracy (%)")

p_rt_camb_task | p_acc_camb_task

ggsave("figures/descriptive/camb_tests.png",
       width = 10, height = 7)
```


```{r}
#| label: plot-cambridge-task-density
#| fig-width: 10
#| fig-height: 5

p_rt_camb_task_dens <- ggplot(rt_pid_camb_task, aes(x = mean, fill = task, colour = task)) +
  geom_density(alpha = 0.7) +
  geom_rug() +
  #facet_wrap(~task) +
  labs(title = "Cambridge RT by task",
       x = "Mean RT (s)", fill = NULL, colour = NULL) +
  theme(legend.position = "bottom")

p_acc_camb_task_dens <- ggplot(acc_pid_camb_task, aes(x = perc, fill = task, colour = task)) +
  geom_density(alpha = 0.7) +
  geom_rug() +
  #facet_wrap(~task) +
  labs(title = "Cambridge accuracy by task",
       x = "Accuracy (%)", fill = NULL, colour = NULL) +
  theme(legend.position = "bottom")

(p_rt_camb_task_dens | p_acc_camb_task_dens) +
  plot_layout(axes = "collect",
              guides = "collect") &
  theme(legend.position = "bottom")

ggsave("figures/descriptive/camb_tests_density.png",
       width = 10, height = 7)
```

## 7.7 Cambridge tests - zoom in on longer average RTs

A few PIDs had long average RTs for the Cambridge tests. 
Let's have a look.

```{r}
# identify high-RT pids
high_rt_pids <- rt_pid_camb |> 
  filter(mean > 20) |> 
  pull(pid) |> 
  unique()

# plot trial-level RT distribution for those pids only
cambridge_tests |>
  filter(pid %in% high_rt_pids) |>
  ggplot(aes(x = rt, fill = task, colour = task)) +
  geom_histogram(binwidth = 5, alpha = 0.7) +
  geom_vline(xintercept = 20, linetype = "dashed", colour = "red") +
  facet_wrap(~pid) +
  labs(title = "Trial-level RT distribution for high-RT participants",
       x = "RT (s)", y = "Count") +
  theme(legend.position = "bottom")

ggsave("figures/descriptive/camb_high_rt_pids.png",
       width = 10, height = 7)

## check how many were extreme
cambridge_tests |>
  filter(pid %in% high_rt_pids) |>
  group_by(pid) |>
  summarise(
    mean_rt  = mean(rt, na.rm = TRUE),
    median_rt = median(rt, na.rm = TRUE),
    n_over_20 = sum(rt > 20, na.rm = TRUE),
    n_over_60 = sum(rt > 60, na.rm = TRUE),
    n_trials  = n()
  )

## check accuracy for those folks
cambridge_tests |>
  filter(pid %in% high_rt_pids) |>
  group_by(pid) |>
  summarise(perc = mean(acc) * 100)
```

---


# 9. Session info

```{r}
#| label: session-info

sessionInfo()
```
