---
title: "Supplementary Materials"
subtitle: "Individual differences in the computational processes that support face recognition"
running-head: "Individual differences and decision models of face memory"
author:
  - name: Raphaël Fournier
    orcid: 0000-0002-5135-4066
    affiliation:
      - ref: 1
  - name: Paul Downing
    affiliation: 
      - ref: 2
  - name: Richard Ramsey
    email: rramsey@ethz.ch
    orcid: 0000-0002-0329-2112
    corresponding: true
    affiliation:
      - ref: 1
      - ref: 3
      - ref: 4
affiliations:
  - id: 1
    name: Neural Control of Movement Laboratory
    department: Department of Health Sciences and Technology, ETH Zurich
  - id: 2
    name: 
    department: School of Psychology and Sports Science, Bangor University
  - id: 3
    name: Social Brain Sciences Laboratory
    department: Department of Humanities, Social and Political Sciences, ETH Zurich
  - id: 4
    name: ETH Zürich
    department: Learning and Instruction Laboratory, Department of Humanities, Social and Political Sciences
keywords: [face perception, memory, familiarity, person knowledge, decision models, evidence accumulation modelling]
bibliography: bibliography.bib
csl: apa.csl
functions: [place]
format:
  preprint-typst:
    output-file: supplementary.pdf
    fontsize: 11pt
    mainfont: Libertinus Serif
editor_options: 
  chunk_output_type: console
knitr:
  opts_chunk:
    dev: "ragg_png"
editor: 
  markdown: 
    wrap: sentence
---
```{r}
#| label: setup
#| include: false

pkgs <- c(
  "tidyverse", "RColorBrewer", "patchwork", "paletteer", 
  "here", "EMC2", "tidybayes", "tinytable", "osfr", "ragg"
)
vapply(
  pkgs, library, logical(1),
  character.only = TRUE, logical.return = TRUE
)

knitr::opts_chunk$set(echo = FALSE)

theme_set(
  theme_bw() +
    theme(text = element_text(size = 15, face = "bold"), 
          title = element_text(size = 15, face = "bold"),
          legend.position = "none")
)
## plot settings ##
## Plot colors
safe_pal <- paletteer_d("rcartocolor::Safe", 4)

myColors_3 <- c(
  "novel"   = safe_pal[1],
  "learned" = safe_pal[2]
)

# Subsets for plots if still needed
myColors_1 <- myColors_3[c("novel", "learned")]

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

{{< pagebreak >}}

## Parameter recovery study



{{< pagebreak >}}

## Priors

These are the priors used for the preregistered LBA model.

```{r}
prior_lba_table <- tibble(
  parameter = c(
    "v_SNovel:lMFALSE", "v_SLearned:lMFALSE",
    "v_SNovel:lMTRUE",  "v_SLearned:lMTRUE",
    "sv_lMTRUE",
    "B_lRNovel", "B_lRLearned",
    "A", "t0"
  ),
  description = c(
    "Drift rate", "Drift rate",
    "Drift rate", "Drift rate",
    "SD drift rate (true accumulator)",
    "Threshold", "Threshold",
    "Start point", "Non-decision time"
  ),
  condition = c(
    "Novel", "Learned",
    "Novel", "Learned",
    "-",
    "-", "-",
    "-", "-"
  ),
  correct = c(
    "False", "False",
    "True",  "True",
    "True",
    "-", "-",
    "-", "-"
  ),
  response = c(
    "-", "-", "-", "-", "-",
    "Novel", "Learned",
    "-", "-"
  ),
  mean = c(
    "0", "0", "1", "1",
    "log(0.5)",
    "log(1)", "log(1)", "log(0.5)", "log(0.3)"
  ),
  sd = c(
    "2", "2", "2", "2",
    "0.5", "0.5", "0.5", "0.3", "0.3"
  )
)

## create the table
prior_lba_table |> 
  tt() |> 
  style_tt(fontsize = 0.85)
```

## Model checks

### Convergence

The chains showed satisfactory mixing as evidenced by the rhat values.

```{r}
#| label: fig-chains-exp1
#| fig-cap: "MCMC chains"
#| fig-width: 8
#| fig-height: 6


```

### Posterior predictive checks

```{r}
#| label: fig-ppreds-exp1
#| fig-cap: "Posterior predictive checks"
#| fig-subcap:
#|   - "CDF"
#|   - "Density"
#| layout-nrow: 2
#| out-height: "40%"

```

### Prior vs posterior plots

```{r}
#| label: fig-prior-post-exp1
#| fig-cap: "Priors vs posterior plots"
#| fig-width: 8
#| fig-height: 6

```


{{< pagebreak >}}

## Point estimate correlations with Cambridge Car Memory score and Cambridge Face Memory scores

```{r}
#| label: load-Cambridge-test
#| include: false
## read in the pid data
cambridge_acc_pid <- read_csv(
  here("data", "cambridge_acc_scores_pid.csv")
)
head(cambridge_acc_pid)
glimpse(cambridge_acc_pid)

## filter and wrangle
## face scores
cam_face <- cambridge_acc_pid |>
  filter(task == "face") |>
  rename(face_recog = n_correct) |>
  mutate(
    face_recog_centred = face_recog - mean(face_recog, na.rm = TRUE)
  ) |>
  select(pid, face_recog, face_recog_centred)

## car scores
cam_car <- cambridge_acc_pid |>
  filter(task == "car") |>
  rename(car_recog = n_correct) |>
  mutate(
    car_recog_centred = car_recog - mean(car_recog, na.rm = TRUE)
  ) |>
  select(pid, car_recog, car_recog_centred)
```

```{r}
#| label: corr-point-estimates
#| include: false
cor_data <- theta_pid_q |>
  select(pid, param, value) |>
  pivot_wider(names_from = param, values_from = value) |>
  left_join(cam_face, by = "pid") |>
  left_join(cam_car, by = "pid") |>
  # derived contrasts
  mutate(
    # drift discrimination (true - false) per condition
    discrim_novel = v_Snovel_lMTRUE - v_Snovel_lMFALSE,
    discrim_learned = v_Slearned_lMTRUE - v_Slearned_lMFALSE,
    # cross-condition drift contrast (novel - learned discrimination)
    discrim_diff = discrim_novel - discrim_learned,
    # threshold contrast
    thresh_diff = B_lRnovel - B_lRlearned
  )


param_cols <- c(
  "v_Snovel_lMTRUE",
  "v_Snovel_lMFALSE",
  "v_Slearned_lMTRUE",
  "v_Slearned_lMFALSE",
  "B_lRnovel",
  "B_lRlearned",
  "discrim_learned",
  "discrim_novel"
)

# pivot to long with both predictors as a variable
cor_long <- cor_data |>
  pivot_longer(
    cols = all_of(param_cols),
    names_to = "param",
    values_to = "estimate"
  ) |>
  pivot_longer(
    cols = c(face_recog_centred, car_recog_centred),
    names_to = "predictor",
    values_to = "score"
  ) |>
  mutate(
    predictor = if_else(str_detect(predictor, "face"), "Face", "Car"),
    predictor = factor(predictor, levels = c("Face", "Car")),
    param = fct_recode(
      param,
      "Drift rate learned (TRUE)" = "v_Slearned_lMTRUE",
      "Drift rate learned (FALSE)" = "v_Slearned_lMFALSE",
      "Drift rate novel (TRUE)" = "v_Snovel_lMTRUE",
      "Drift rate novel (FALSE)" = "v_Snovel_lMFALSE",
      "Threshold novel" = "B_lRnovel",
      "Threshold learned" = "B_lRlearned",
      "Drift rate learned (TRUE-FALSE)" = "discrim_learned",
      "Drift rate novel (TRUE-FALSE)" = "discrim_novel"
    )
  )

# compute r per param x predictor
cor_long_thresh <- cor_long |>
  filter(param %in% c("Threshold learned", "Threshold novel"))

cor_labels_thresh <- cor_long_thresh |>
  group_by(param, predictor) |>
  summarise(
    r = cor(estimate, score, use = "complete.obs"),
    .groups = "drop"
  ) |>
  mutate(
    label = sprintf("r = %.2f", r),
    vjust = if_else(predictor == "Face", 1.5, 3.0)
  )

cor_long_drift <- cor_long |>
  filter(
    param %in%
      c(
        "Drift rate learned (TRUE)",
        "Drift rate learned (FALSE)",
        "Drift rate learned (TRUE-FALSE)",
        "Drift rate novel (TRUE)",
        "Drift rate novel (FALSE)",
        "Drift rate novel (TRUE-FALSE)"
      )
  ) |>
  mutate(
    param = fct_relevel(
      param,
      "Drift rate learned (TRUE)",
      "Drift rate learned (FALSE)",
      "Drift rate learned (TRUE-FALSE)",
      "Drift rate novel (TRUE)",
      "Drift rate novel (FALSE)",
      "Drift rate novel (TRUE-FALSE)"
    )
  )

cor_labels_drift <- cor_long_drift |>
  group_by(param, predictor) |>
  summarise(
    r = cor(estimate, score, use = "complete.obs"),
    .groups = "drop"
  ) |>
  mutate(
    label = sprintf("r = %.2f", r),
    vjust = if_else(predictor == "Face", 1.5, 3.0),
    param = fct_relevel(
      param,
      "Drift rate learned (TRUE)",
      "Drift rate learned (FALSE)",
      "Drift rate learned (TRUE-FALSE)",
      "Drift rate novel (TRUE)",
      "Drift rate novel (FALSE)",
      "Drift rate novel (TRUE-FALSE)"
    )
  )
```

```{r}
#| label: prepare-plot-corr-point-estimate
#| include: false
# plot thresh
p_raw_params_thresh <- ggplot(
  cor_long_thresh,
  aes(x = score, y = estimate, colour = predictor, fill = predictor)
) +
  geom_point(alpha = 0.3, size = 2) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.3) +
  geom_text(
    data = cor_labels_thresh,
    aes(label = label, vjust = vjust, colour = predictor),
    x = -Inf,
    y = Inf,
    hjust = -0.1,
    size = 5,
    fontface = "bold",
    show.legend = FALSE
  ) +
  facet_wrap(~param, scales = "free_y") +
  force_panelsizes(cols = c(1, 1), rows = 0.5) +
  scale_colour_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  scale_fill_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  labs(
    x = "Recognition score (centred)",
    y = "Parameter estimate (median)",
    colour = NULL,
    fill = NULL
  ) +
  theme(legend.position = "bottom")


# plot drift
p_raw_params_drift <- ggplot(
  cor_long_drift,
  aes(x = score, y = estimate, colour = predictor, fill = predictor)
) +
  geom_point(alpha = 0.3, size = 2) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.3) +
  geom_text(
    data = cor_labels_drift,
    aes(label = label, vjust = vjust, colour = predictor),
    x = -Inf,
    y = Inf,
    hjust = -0.1,
    size = 5,
    fontface = "bold",
    show.legend = FALSE
  ) +
  facet_wrap(~param, scales = "free_y", ncol = 3) +
  scale_colour_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  scale_fill_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  labs(
    x = "Recognition score (centred)",
    y = "Parameter estimate (median)",
    colour = NULL,
    fill = NULL
  ) +
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-corr-point-estimates
#| fig-cap: "Point estimates correlations with recognition scores. In (b), drift rate (TRUE) reflects the accumulation of evidence when the response is correct and drift rate (FALSE) the accumulation of evidence when the response is incorrect. Drift rate (TRUE-FALSE) represent the signal-to-noise ratio provided by the stimulus."
#| fig-subcap: 
#|  - "Median threshold correlations with recognition ability"
#|  - "Median drift rate correlations with recognition ability"
#| layout-ncol: 1
#| warning: false
#| fig-width: 10
#| fig-height: 7
p_raw_params_thresh
p_raw_params_drift
```

{{< pagebreak >}}

## Plausible value correlations with Cambridge Car Memory scores and Cambridge Face Memory scores

```{r}
#| label: plausible-value-wrangle
#| include: false
param_cols <- c(
  "v_Snovel_lMTRUE",
  "v_Snovel_lMFALSE",
  "v_Slearned_lMTRUE",
  "v_Slearned_lMFALSE",
  "sv_lMTRUE",
  "B_lRnovel",
  "B_lRlearned",
  "A",
  "t0"
)

all_prereg_params <- c(
  param_cols,
  "discrim_novel",
  "discrim_learned"
)

param_labels <- c(
  B_lRnovel = "Threshold: Novel",
  B_lRlearned = "Threshold: Learned",
  v_Snovel_lMTRUE = "Drift: Novel (TRUE)",
  v_Snovel_lMFALSE = "Drift: Novel (FALSE)",
  discrim_novel = "Drift: Novel (TRUE-FALSE)",
  v_Slearned_lMTRUE = "Drift: Learned (TRUE)",
  v_Slearned_lMFALSE = "Drift: Learned (FALSE)",
  discrim_learned = "Drift: Learned (TRUE-FALSE)"
)

# join theta_pid (full posterior) with cambridge scores
# theta_pid has 300 pids x 9 params x 15000 draws
pv_data <- theta_pid |>
  pivot_wider(names_from = param, values_from = value) |>
  mutate(
    discrim_novel = v_Snovel_lMTRUE - v_Snovel_lMFALSE,
    discrim_learned = v_Slearned_lMTRUE - v_Slearned_lMFALSE
  ) |>
  left_join(
    cam_face |> select(pid, face_recog_centred),
    by = "pid"
  ) |>
  left_join(
    cam_car |> select(pid, car_recog_centred),
    by = "pid"
  ) |>
  pivot_longer(
    cols = all_of(all_prereg_params),
    names_to = "param",
    values_to = "estimate"
  )

# compute correlation for each draw x param x predictor
# this will take a moment with 15000 draws
pv_cors <- pv_data |>
  group_by(.draw, param) |>
  summarise(
    r_face = cor(estimate, face_recog_centred, use = "complete.obs"),
    r_car = cor(estimate, car_recog_centred, use = "complete.obs"),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = c(r_face, r_car),
    names_to = "predictor",
    values_to = "r"
  ) |>
  mutate(predictor = if_else(str_detect(predictor, "_face"), "Face", "Car"))
pv_summary <- pv_cors |>
  group_by(param, predictor) |>
  median_qi(r, .width = c(0.95)) |>
  mutate(label = param_labels[param]) |>
  print(n = Inf)
```

```{r}
#| label: create-plausible-value-corr-plots-thresh
#| include: false
# left panel — novel and learned
pv_thresh_main <- pv_cors |>
  filter(param %in% c("B_lRnovel", "B_lRlearned")) |>
  mutate(
    response = if_else(param == "B_lRnovel", "Novel", "Learned"),
    response = factor(response, levels = c("Learned", "Novel"))
  )

p_thresh_left <- ggplot(
  pv_thresh_main,
  aes(x = r, y = predictor, fill = response, colour = response)
) +
  stat_halfeye(alpha = 0.7, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = myColors_4) +
  scale_colour_manual(values = myColors_4) +
  labs(
    x = "Plausible r value",
    y = NULL,
    fill = NULL,
    colour = NULL
  ) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 5, face = "bold"),
    strip.text = element_text(size = 5, face = "bold"),
    axis.text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12, face = "bold")
  )
```

```{r}
#| label: create-plausible-value-corr-plots-drift
#| include: false
# first panel - drift rate true and false
pv_drift_true_false <- pv_cors |>
  filter(
    param %in%
      c(
        "v_Slearned_lMTRUE",
        "v_Slearned_lMFALSE",
        "v_Snovel_lMTRUE",
        "v_Snovel_lMFALSE"
      )
  ) |>
  mutate(
    condition = if_else(
      param == "v_Slearned_lMTRUE",
      "Learned",
      if_else(param == "v_Slearned_lMFALSE", "Learned", "Novel")
    ),
    accuracy = if_else(
      param == "v_Slearned_lMTRUE",
      "TRUE",
      if_else(param == "v_Snovel_lMTRUE", "TRUE", "FALSE")
    ),
    accuracy = factor(accuracy, levels = c("TRUE", "FALSE"))
  )

p_drift_first <- ggplot(
  pv_drift_true_false,
  aes(x = r, y = predictor, fill = accuracy, colour = accuracy)
) +
  stat_halfeye(alpha = 0.7, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = myColors_acc_2) +
  scale_colour_manual(values = myColors_acc_2) +
  labs(
    x = "Plausible r value",
    y = NULL,
    fill = NULL,
    colour = NULL
  ) +
  facet_wrap(~ condition) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 7, face = "bold"),
    strip.text = element_text(size = 10, face = "bold"),
    axis.text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12, face = "bold")
  )

# middle panel — discrimination (true - false) per condition
pv_drift_discrim <- pv_cors |>
  filter(param %in% c("discrim_novel", "discrim_learned")) |>
  mutate(
    condition = if_else(param == "discrim_novel", "Novel", "Learned"),
    condition = factor(condition, levels = c("Learned", "Novel"))
  )

p_drift_mid <- ggplot(
  pv_drift_discrim,
  aes(x = r, y = predictor, fill = condition, colour = condition)
) +
  stat_halfeye(alpha = 0.7, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = myColors_4) +
  scale_colour_manual(values = myColors_4) +
  labs(
    x = "Plausible r value",
    y = NULL,
    fill = NULL,
    colour = NULL
  ) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 7, face = "bold"),
    strip.text = element_text(size = 7, face = "bold"),
    axis.text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12, face = "bold")
  )
```

```{r}
#| label: tbl-plausible-cor
#| tbl-cap: "Median r and 95% quantile intervals of the plausible value correlations between LBA parameters and recogntion ability scores."
#| include: true
#| tbl-height: 10
#| tbl-width: 15
table_data <- pv_summary |>
  filter(!is.na(label)) |>
  # no longer drop true/false rows
  mutate(
    r_fmt = sprintf("%.2f", r),
    ci_fmt = sprintf("[%.2f, %.2f]", .lower, .upper),
    color = case_when(
      .lower > 0 & .upper > 0 ~ "blue",
      .lower < 0 & .upper < 0 ~ "red",
      TRUE ~ "black"
    )
  ) |>
  select(label, predictor, r_fmt, ci_fmt, color) |>
  pivot_wider(
    names_from = predictor,
    values_from = c(r_fmt, ci_fmt, color),
    names_glue = "{predictor}_{.value}"
  ) |>
  separate(
    label,
    into = c("Parameter", "Condition"),
    sep = ": ",
    fill = "right"
  ) |>
  mutate(
    # extract accuracy before cleaning condition
    Accuracy = case_when(
      grepl("\\(TRUE\\)$", Condition, ignore.case = TRUE) ~ "TRUE",
      grepl("\\(FALSE\\)$", Condition, ignore.case = TRUE) ~ "FALSE",
      grepl("TRUE-FALSE", Condition, ignore.case = TRUE) ~ "TRUE-FALSE",
      TRUE ~ "-"
    ),
    # clean up condition labels
    Condition = case_when(
      Condition == "Learned (TRUE)" ~ "Learned",
      Condition == "Learned (FALSE)" ~ "Learned",
      Condition == "Novel (TRUE)" ~ "Novel",
      Condition == "Novel (FALSE)" ~ "Novel",
      Condition == "Learned (TRUE-FALSE)" ~ "Learned",
      Condition == "Novel (TRUE-FALSE)" ~ "Novel",
      TRUE ~ Condition
    ),
    # set ordering
    Parameter = factor(Parameter, levels = c("Threshold", "Drift")),
    Condition = factor(
      Condition,
      levels = c("Learned", "Novel")
    ),
    Accuracy = factor(Accuracy, levels = c("TRUE", "FALSE", "TRUE-FALSE", "-"))
  ) |>
  arrange(Parameter, Condition, Accuracy)


# rename columns to clean display names first
table_data_tt <- table_data |>
  rename(
    `median r_Face` = Face_r_fmt,
    `95% QI_Face` = Face_ci_fmt,
    `median r_Car` = Car_r_fmt,
    `95% QI_Car` = Car_ci_fmt
  ) |>
  select(
    Parameter,
    Condition,
    Accuracy,
    `median r_Face`,
    `95% QI_Face`,
    `median r_Car`,
    `95% QI_Car`,
    Face_color,
    Car_color
  )
table_data_tt <- as.data.frame(table_data_tt)

tt_table <- tt(
  table_data_tt |> select(-Face_color, -Car_color),
  notes = "Note. Blue = 95% QI entirely above zero; red = 95% QI entirely below zero."
) |>
  group_tt(j = list("Face" = 4:5, "Car" = 6:7)) |>
  style_tt(i = 0, bold = TRUE) |>
  style_tt(i = 1:nrow(table_data_tt), fontsize = 0.7) |>
  style_tt(i = 0, fontsize = 0.8) |>
  style_tt(i = -1, bold = TRUE, fontsize = 0.8) |>
  style_tt(
    i = which(table_data_tt$Face_color == "blue"),
    j = 4:5,
    color = "blue"
  ) |>
  style_tt(
    i = which(table_data_tt$Face_color == "red"),
    j = 4:5,
    color = "red"
  ) |>
  style_tt(
    i = which(table_data_tt$Car_color == "blue"),
    j = 6:7,
    color = "blue"
  ) |>
  style_tt(
    i = which(table_data_tt$Car_color == "red"),
    j = 6:7,
    color = "red"
  )

tt_table
```


```{r}

```


## Bayesian linear regression models for threshold parameter

```{r}

```



