---
title: "Supplementary Materials"
subtitle: "Individual differences in the computational processes that support face recognition"
running-head: "Individual differences and decision models of face memory"
author:
  - name: Raphaël Fournier
    orcid: 0000-0002-5135-4066
    affiliation:
      - ref: 1
  - name: Paul Downing
    affiliation: 
      - ref: 2
  - name: Richard Ramsey
    email: rramsey@ethz.ch
    orcid: 0000-0002-0329-2112
    corresponding: true
    affiliation:
      - ref: 1
      - ref: 3
      - ref: 4
affiliations:
  - id: 1
    name: Neural Control of Movement Laboratory
    department: Department of Health Sciences and Technology, ETH Zurich
  - id: 2
    name: 
    department: School of Psychology and Sports Science, Bangor University
  - id: 3
    name: Social Brain Sciences Laboratory
    department: Department of Humanities, Social and Political Sciences, ETH Zurich
  - id: 4
    name: ETH Zürich
    department: Learning and Instruction Laboratory, Department of Humanities, Social and Political Sciences
keywords: [face perception, memory, familiarity, person knowledge, decision models, evidence accumulation modelling]
functions: [place]
format:
  preprint-typst:
    output-file: supplementary.pdf
    fontsize: 11pt
    mainfont: Libertinus Serif
editor_options: 
  chunk_output_type: console
knitr:
  opts_chunk:
    dev: "ragg_png"
editor: 
  markdown: 
    wrap: sentence
---
```{r}
#| label: setup
#| include: false
pkgs <- c(
  "tidyverse",
  "RColorBrewer",
  "patchwork",
  "paletteer",
  "here",
  "EMC2",
  "tidybayes",
  "tinytable",
  "osfr",
  "svglite",
  "gt",
  "brms",
  "rstantool",
  "corrplot",
  "ggdist",
  "parallel",
  "future",
  "ggplot2",
  "ggh4x"
)
vapply(
  pkgs,
  library,
  logical(1),
  character.only = TRUE,
  logical.return = TRUE
)

knitr::opts_chunk$set(echo = FALSE)

theme_set(
  theme_bw() +
    theme(
      text = element_text(size = 15, face = "bold"),
      title = element_text(size = 15, face = "bold"),
      legend.position = "none"
    )
)
## plot settings ##
## Plot colors
safe_pal <- paletteer_d("rcartocolor::Safe", 4)
dark_pal <- paletteer_d("RColorBrewer::Dark2", 2)

myColors_3 <- c(
  "novel" = safe_pal[1],
  "learned" = safe_pal[2]
)
myColors_4 <- c(
  "Novel" = safe_pal[1],
  "Learned" = safe_pal[2]
)

#Accuracy colors
myColors_acc <- c(
  "true" = dark_pal[1],
  "false" = dark_pal[2]
)
myColors_acc_2 <- c(
  "TRUE" = dark_pal[1],
  "FALSE" = dark_pal[2]
)
# Subsets for single-experiment plots if still needed
myColors_1 <- myColors_3[c("novel", "learned")]


## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

{{< pagebreak >}}

## Parameter recovery study



{{< pagebreak >}}

## Priors

These are the priors used for the preregistered LBA model.

```{r}
prior_lba_table <- tibble(
  parameter = c(
    "v_SNovel:lMFALSE", "v_SLearned:lMFALSE",
    "v_SNovel:lMTRUE",  "v_SLearned:lMTRUE",
    "sv_lMTRUE",
    "B_lRNovel", "B_lRLearned",
    "A", "t0"
  ),
  description = c(
    "Drift rate", "Drift rate",
    "Drift rate", "Drift rate",
    "SD drift rate (true accumulator)",
    "Threshold", "Threshold",
    "Start point", "Non-decision time"
  ),
  condition = c(
    "Novel", "Learned",
    "Novel", "Learned",
    "-",
    "-", "-",
    "-", "-"
  ),
  correct = c(
    "False", "False",
    "True",  "True",
    "True",
    "-", "-",
    "-", "-"
  ),
  response = c(
    "-", "-", "-", "-", "-",
    "Novel", "Learned",
    "-", "-"
  ),
  mean = c(
    "0", "0", "1", "1",
    "log(0.5)",
    "log(1)", "log(1)", "log(0.5)", "log(0.3)"
  ),
  sd = c(
    "2", "2", "2", "2",
    "0.5", "0.5", "0.5", "0.3", "0.3"
  )
)

## create the table
prior_lba_table |> 
  tt() |> 
  style_tt(fontsize = 0.85)
```

## Model checks

### Convergence

The chains showed satisfactory mixing as evidenced by the rhat values.

```{r}
#| label: fig-chains-exp1
#| fig-cap: "MCMC chains"
#| fig-width: 8
#| fig-height: 6


```

### Posterior predictive checks

```{r}
#| label: fig-ppreds-exp1
#| fig-cap: "Posterior predictive checks"
#| fig-subcap:
#|   - "CDF"
#|   - "Density"
#| layout-nrow: 2
#| out-height: "40%"

```

### Prior vs posterior plots

```{r}
#| label: fig-prior-post-exp1
#| fig-cap: "Priors vs posterior plots"
#| fig-width: 8
#| fig-height: 6

```

{{< pagebreak >}}

## Subject-level drift rates and thresholds

To visualize the individual variability of our parameters of interest, subject-level posterior distributions for threshold (Learned and Novel) and drift rate (Learned and novel) can be found in @fig-ind-param.

```{r}
#| label: load-lba-data
#| include: false
theta_group <- read_csv(here("data/theta_group.csv"))
theta_group_q <- read_csv(here("data/theta_group_q.csv"))
theta_pid <- read_csv(here("data/theta_pid.csv"))
theta_pid_q <- read_csv(here("data/theta_pid_q.csv"))
```

```{r}
#| label: wrangle-threshold-drift rate
#| include: false

#Wrangle threshold
thresh_group <- theta_group |>
  filter(str_detect(param, "B_")) |>
  mutate(
    response = if_else(str_detect(param, "lRlearned"), "Learned", "Novel"),
    response = factor(response, levels = c("Novel", "Learned"))
  )
post_th <- theta_pid |>
  filter(str_detect(param, "B_")) |>
  mutate(
    response = if_else(str_detect(param, "lRlearned"), "Learned", "Novel"),
    response = factor(response, levels = c("Novel", "Learned"))
  ) |>
  left_join(thresh_group, by = c("response",".draw")) |>
  select(pid, .draw, response, value.x, value.y) |>
  rename(subject = "value.x", group = "value.y") 
  
#Wrangle drift rate
drift_group <- theta_group |>
  filter(str_detect(param, "v_S")) |>
  mutate(
    condition = if_else(str_detect(param, "novel"), "Novel", "Learned"),
    accuracy = if_else(str_detect(param, "TRUE"), "true", "false"),
    condition = factor(condition, levels = c("Novel", "Learned")),
    accuracy = factor(accuracy, levels = c("false", "true"))
  )

drift_pid <- theta_pid |>
  filter(str_detect(param, "v_S")) |>
  mutate(
    condition = if_else(str_detect(param, "novel"), "Novel", "Learned"),
    accuracy = if_else(str_detect(param, "TRUE"), "true", "false"),
    condition = factor(condition, levels = c("Novel", "Learned")),
    accuracy = factor(accuracy, levels = c("false", "true"))
  )

# true - false per condition
drift_acc_diff_group <- drift_group |>
  pivot_wider(
    id_cols = c(.draw, condition),
    names_from = accuracy,
    values_from = value
  ) |>
  mutate(diff = true - false)

drift_acc_diff_pid <- drift_pid |>
  pivot_wider(
    id_cols = c(pid, .draw, condition),
    names_from = accuracy,
    values_from = value
  ) |>
  mutate(diff = true - false)

# novel - learned contrast
drift_cond_diff_group <- drift_acc_diff_group |>
  pivot_wider(id_cols = .draw, names_from = condition, values_from = diff) |>
  mutate(diff = Novel -Learned)

drift_cond_diff_pid <- drift_acc_diff_pid |>
  pivot_wider(
    id_cols = c(pid, .draw),
    names_from = condition,
    values_from = diff
  ) |> 
  left_join(drift_cond_diff_group, by = ".draw") |>
  select(pid, .draw, Novel.x, Novel.y, Learned.x, Learned.y) |>
  rename(Novel_subject = "Novel.x", Novel_group = "Novel.y", Learned_subject = "Learned.x", Learned_group = "Learned.y") 

#Wide -> long: one row per pid x draw x response
post_dr <- drift_cond_diff_pid |>
  pivot_longer(
    cols      = -c(pid, .draw),
    names_to  = c("stimulus", ".value"),
    names_sep = "_"
  ) |>
  mutate(stimulus = factor(stimulus, levels = c("Novel", "Learned"))) 
```

```{r}
#| label: combine-thresh-drift
#| include: false

# post_thr = threshold data frame  (pid, .draw, response, subject, group)
# post_dr  = drift rate data frame (pid, .draw, stimulus, subject, group)

# 0. Harmonise and combine
panel_levels <- c("Threshold Learned", "Threshold Novel",
                  "Drift rate Learned", "Drift rate Novel")

post_all <- bind_rows(
  post_th |> rename(condition = response) |> mutate(parameter = "Threshold"),
  post_dr  |> rename(condition = stimulus) |> mutate(parameter = "Drift rate")
) |>
  mutate(panel = factor(paste(parameter, condition), levels = panel_levels))

# 1. Group-level draws: one row per draw x panel
group_draws <- post_all |>
  distinct(.draw, panel, group) |>
  transmute(.draw, panel, label = "Group", value = group, level = "Group")

# 2. Participant-level draws
pid_draws <- post_all |>
  transmute(.draw, panel, label = as.character(pid), value = subject, level = "Participant")

# 3. Order pids by posterior median within each panel (ascending, because the
#    first level is drawn at the bottom), with the group row last = top
key_levels <- pid_draws |>
  group_by(panel, label) |>
  summarise(med = median(value), .groups = "drop") |>
  arrange(panel, med) |>
  mutate(key = paste(label, panel, sep = "___")) |>
  group_split(panel) |>
  map(\(d) c(d$key, paste("Group", d$panel[1], sep = "___"))) |>
  unlist()

plot_df <- bind_rows(group_draws, pid_draws) |>
  mutate(key = factor(paste(label, panel, sep = "___"), levels = key_levels))

# Separator line between the pids and the group row, computed per panel
sep_df <- pid_draws |>
  group_by(panel) |>
  summarise(y = n_distinct(label) + 0.5)

# 4. Plot
indiv_param_plot <- ggplot(plot_df, aes(x = value, y = key, fill = level)) +
  stat_halfeye(.width = 0.95, point_interval = median_qi) +
  geom_hline(data = sep_df, aes(yintercept = y), linetype = "dotted") +
  scale_y_discrete(labels = \(x) sub("___.*", "", x)) +   # show pid / "Group" only
  scale_fill_manual(values = c(Group = "#F8766D", Participant = "grey70")) +
  facet_wrap(~ panel, ncol = 4, scales = "free") +
  labs(x = "Subject-level posterior estimates", y = "", fill = NULL) +
  theme_bw() +
  theme(axis.text.y = element_blank())
```


```{r}
#| label: fig-ind-param
#| fig-cap: "Subject-level posterior distribution of the threshold parameter (Learned and Novel) and the drift rate parameter (Learned and Novel). The top distribution visualize the group-level distribution. Black points denotate the median of the distribtuiona and error bars the 95% QI intervals."
#| include: true
#| fig-width: 10
#| fig-height: 15
indiv_param_plot
```

{{< pagebreak >}}

## Point estimate correlations with Cambridge Car Memory score and Cambridge Face Memory scores


```{r}
#| label: load-Cambridge-test
#| include: false
## read in the pid data
cambridge_acc_pid <- read_csv(
  here("data", "cambridge_acc_scores_pid.csv")
)
head(cambridge_acc_pid)
glimpse(cambridge_acc_pid)

## filter and wrangle
## face scores
cam_face <- cambridge_acc_pid |>
  filter(task == "face") |>
  rename(face_recog = n_correct) |>
  mutate(
    face_recog_centred = face_recog - mean(face_recog, na.rm = TRUE)
  ) |>
  select(pid, face_recog, face_recog_centred)

## car scores
cam_car <- cambridge_acc_pid |>
  filter(task == "car") |>
  rename(car_recog = n_correct) |>
  mutate(
    car_recog_centred = car_recog - mean(car_recog, na.rm = TRUE)
  ) |>
  select(pid, car_recog, car_recog_centred)
```

```{r}
#| label: corr-point-estimates
#| include: false
cor_data <- theta_pid_q |>
  select(pid, param, value) |>
  pivot_wider(names_from = param, values_from = value) |>
  left_join(cam_face, by = "pid") |>
  left_join(cam_car, by = "pid") |>
  # derived contrasts
  mutate(
    # drift discrimination (true - false) per condition
    discrim_novel = v_Snovel_lMTRUE - v_Snovel_lMFALSE,
    discrim_learned = v_Slearned_lMTRUE - v_Slearned_lMFALSE,
    # cross-condition drift contrast (novel - learned discrimination)
    discrim_diff = discrim_novel - discrim_learned,
    # threshold contrast
    thresh_diff = B_lRnovel - B_lRlearned
  )


param_cols <- c(
  "v_Snovel_lMTRUE",
  "v_Snovel_lMFALSE",
  "v_Slearned_lMTRUE",
  "v_Slearned_lMFALSE",
  "B_lRnovel",
  "B_lRlearned",
  "discrim_learned",
  "discrim_novel"
)

# pivot to long with both predictors as a variable
cor_long <- cor_data |>
  pivot_longer(
    cols = all_of(param_cols),
    names_to = "param",
    values_to = "estimate"
  ) |>
  pivot_longer(
    cols = c(face_recog_centred, car_recog_centred),
    names_to = "predictor",
    values_to = "score"
  ) |>
  mutate(
    predictor = if_else(str_detect(predictor, "face"), "Face", "Car"),
    predictor = factor(predictor, levels = c("Face", "Car")),
    param = fct_recode(
      param,
      "Drift rate learned (TRUE)" = "v_Slearned_lMTRUE",
      "Drift rate learned (FALSE)" = "v_Slearned_lMFALSE",
      "Drift rate novel (TRUE)" = "v_Snovel_lMTRUE",
      "Drift rate novel (FALSE)" = "v_Snovel_lMFALSE",
      "Threshold novel" = "B_lRnovel",
      "Threshold learned" = "B_lRlearned",
      "Drift rate learned (TRUE-FALSE)" = "discrim_learned",
      "Drift rate novel (TRUE-FALSE)" = "discrim_novel"
    )
  )

# compute r per param x predictor
cor_long_thresh <- cor_long |>
  filter(param %in% c("Threshold learned", "Threshold novel"))

cor_labels_thresh <- cor_long_thresh |>
  group_by(param, predictor) |>
  summarise(
    r = cor(estimate, score, use = "complete.obs"),
    .groups = "drop"
  ) |>
  mutate(
    label = sprintf("r = %.2f", r),
    vjust = if_else(predictor == "Face", 1.5, 3.0)
  )

cor_long_drift <- cor_long |>
  filter(
    param %in%
      c(
        "Drift rate learned (TRUE)",
        "Drift rate learned (FALSE)",
        "Drift rate learned (TRUE-FALSE)",
        "Drift rate novel (TRUE)",
        "Drift rate novel (FALSE)",
        "Drift rate novel (TRUE-FALSE)"
      )
  ) |>
  mutate(
    param = fct_relevel(
      param,
      "Drift rate learned (TRUE)",
      "Drift rate learned (FALSE)",
      "Drift rate learned (TRUE-FALSE)",
      "Drift rate novel (TRUE)",
      "Drift rate novel (FALSE)",
      "Drift rate novel (TRUE-FALSE)"
    )
  )

cor_labels_drift <- cor_long_drift |>
  group_by(param, predictor) |>
  summarise(
    r = cor(estimate, score, use = "complete.obs"),
    .groups = "drop"
  ) |>
  mutate(
    label = sprintf("r = %.2f", r),
    vjust = if_else(predictor == "Face", 1.5, 3.0),
    param = fct_relevel(
      param,
      "Drift rate learned (TRUE)",
      "Drift rate learned (FALSE)",
      "Drift rate learned (TRUE-FALSE)",
      "Drift rate novel (TRUE)",
      "Drift rate novel (FALSE)",
      "Drift rate novel (TRUE-FALSE)"
    )
  )
```

```{r}
#| label: prepare-plot-corr-point-estimate
#| include: false
# plot thresh
p_raw_params_thresh <- ggplot(
  cor_long_thresh,
  aes(x = score, y = estimate, colour = predictor, fill = predictor)
) +
  geom_point(alpha = 0.3, size = 2) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.3) +
  geom_text(
    data = cor_labels_thresh,
    aes(label = label, vjust = vjust, colour = predictor),
    x = -Inf,
    y = Inf,
    hjust = -0.1,
    size = 5,
    fontface = "bold",
    show.legend = FALSE
  ) +
  facet_wrap(~param, scales = "free_y") +
  scale_colour_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  scale_fill_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  labs(
    x = "Recognition score (centred)",
    y = "Parameter estimate (median)",
    colour = NULL,
    fill = NULL
  ) +
  theme(legend.position = "bottom")


# plot drift
p_raw_params_drift <- ggplot(
  cor_long_drift,
  aes(x = score, y = estimate, colour = predictor, fill = predictor)
) +
  geom_point(alpha = 0.3, size = 2) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.3) +
  geom_text(
    data = cor_labels_drift,
    aes(label = label, vjust = vjust, colour = predictor),
    x = -Inf,
    y = Inf,
    hjust = -0.1,
    size = 5,
    fontface = "bold",
    show.legend = FALSE
  ) +
  facet_wrap(~param, scales = "free_y", ncol = 3) +
  scale_colour_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  scale_fill_manual(values = c("Face" = safe_pal[1], "Car" = safe_pal[2])) +
  labs(
    x = "Recognition score (centred)",
    y = "Parameter estimate (median)",
    colour = NULL,
    fill = NULL
  ) +
  theme(legend.position = "bottom")
```



```{r}
#| label: fig-corr-point-estimates
#| fig-cap: "Point estimates correlations with recognition scores. In (b), drift rate (TRUE) reflects the accumulation of evidence when the response is correct and drift rate (FALSE) the accumulation of evidence when the response is incorrect. Drift rate (TRUE-FALSE) represent the signal-to-noise ratio provided by the stimulus."
#| fig-subcap: 
#|  - "Median threshold correlations with recognition ability"
#|  - "Median drift rate correlations with recognition ability"
#| layout-ncol: 1
#| warning: false
#| fig-width: 10
#| fig-height: 7
p_raw_params_thresh
p_raw_params_drift
```

{{< pagebreak >}}

## Plausible value correlations with Cambridge Car Memory scores and Cambridge Face Memory scores

We calculated the posterior correlations distributions between car recognition ability and our parameters of interest. Correlations with threshold can be found in @fig-plausible-corr-thresh and correlations with drift rate in @fig-plausible-corr-drift. @tbl-plausible-cor summarize these distributions. The posterior correlations with face recognition ability were added to facilitate comparisons between the two memory scores. 

```{r}
#| label: plausible-value-wrangle
#| include: false
param_cols <- c(
  "v_Snovel_lMTRUE",
  "v_Snovel_lMFALSE",
  "v_Slearned_lMTRUE",
  "v_Slearned_lMFALSE",
  "sv_lMTRUE",
  "B_lRnovel",
  "B_lRlearned",
  "A",
  "t0"
)

all_prereg_params <- c(
  param_cols,
  "discrim_novel",
  "discrim_learned"
)

param_labels <- c(
  B_lRnovel = "Threshold: Novel",
  B_lRlearned = "Threshold: Learned",
  v_Snovel_lMTRUE = "Drift: Novel (TRUE)",
  v_Snovel_lMFALSE = "Drift: Novel (FALSE)",
  discrim_novel = "Drift: Novel (TRUE-FALSE)",
  v_Slearned_lMTRUE = "Drift: Learned (TRUE)",
  v_Slearned_lMFALSE = "Drift: Learned (FALSE)",
  discrim_learned = "Drift: Learned (TRUE-FALSE)"
)

# join theta_pid (full posterior) with cambridge scores
# theta_pid has 300 pids x 9 params x 15000 draws
pv_data <- theta_pid |>
  pivot_wider(names_from = param, values_from = value) |>
  mutate(
    discrim_novel = v_Snovel_lMTRUE - v_Snovel_lMFALSE,
    discrim_learned = v_Slearned_lMTRUE - v_Slearned_lMFALSE
  ) |>
  left_join(
    cam_face |> select(pid, face_recog_centred),
    by = "pid"
  ) |>
  left_join(
    cam_car |> select(pid, car_recog_centred),
    by = "pid"
  ) |>
  pivot_longer(
    cols = all_of(all_prereg_params),
    names_to = "param",
    values_to = "estimate"
  )

# compute correlation for each draw x param x predictor
# this will take a moment with 15000 draws
pv_cors <- pv_data |>
  group_by(.draw, param) |>
  summarise(
    r_face = cor(estimate, face_recog_centred, use = "complete.obs"),
    r_car = cor(estimate, car_recog_centred, use = "complete.obs"),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = c(r_face, r_car),
    names_to = "predictor",
    values_to = "r"
  ) |>
  mutate(predictor = if_else(str_detect(predictor, "_face"), "Face", "Car"))
pv_summary <- pv_cors |>
  group_by(param, predictor) |>
  median_qi(r, .width = c(0.95)) |>
  mutate(label = param_labels[param]) |>
  print(n = Inf)
```

```{r}
#| label: create-plausible-value-corr-plots-thresh
#| include: false
# left panel — novel and learned
pv_thresh_main <- pv_cors |>
  filter(param %in% c("B_lRnovel", "B_lRlearned")) |>
  mutate(
    response = if_else(param == "B_lRnovel", "Novel", "Learned"),
    response = factor(response, levels = c("Learned", "Novel"))
  )

p_thresh_left <- ggplot(
  pv_thresh_main,
  aes(x = r, y = predictor, fill = response, colour = response)
) +
  stat_halfeye(alpha = 0.7, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = myColors_4) +
  scale_colour_manual(values = myColors_4) +
  labs(
    x = "Plausible r value",
    y = NULL,
    fill = NULL,
    colour = NULL
  ) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 5, face = "bold"),
    strip.text = element_text(size = 5, face = "bold"),
    axis.text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12, face = "bold")
  )
```

```{r}
#| label: create-plausible-value-corr-plots-drift
#| include: false
# first panel - drift rate true and false
pv_drift_true_false <- pv_cors |>
  filter(
    param %in%
      c(
        "v_Slearned_lMTRUE",
        "v_Slearned_lMFALSE",
        "v_Snovel_lMTRUE",
        "v_Snovel_lMFALSE"
      )
  ) |>
  mutate(
    condition = if_else(
      param == "v_Slearned_lMTRUE",
      "Learned",
      if_else(param == "v_Slearned_lMFALSE", "Learned", "Novel")
    ),
    accuracy = if_else(
      param == "v_Slearned_lMTRUE",
      "TRUE",
      if_else(param == "v_Snovel_lMTRUE", "TRUE", "FALSE")
    ),
    accuracy = factor(accuracy, levels = c("TRUE", "FALSE"))
  )

p_drift_first <- ggplot(
  pv_drift_true_false,
  aes(x = r, y = predictor, fill = accuracy, colour = accuracy)
) +
  stat_halfeye(alpha = 0.7, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = myColors_acc_2) +
  scale_colour_manual(values = myColors_acc_2) +
  labs(
    x = "Plausible r value",
    y = NULL,
    fill = NULL,
    colour = NULL
  ) +
  facet_wrap(~ condition) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 7, face = "bold"),
    strip.text = element_text(size = 10, face = "bold"),
    axis.text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12, face = "bold")
  )

# middle panel — discrimination (true - false) per condition
pv_drift_discrim <- pv_cors |>
  filter(param %in% c("discrim_novel", "discrim_learned")) |>
  mutate(
    condition = if_else(param == "discrim_novel", "Novel", "Learned"),
    condition = factor(condition, levels = c("Learned", "Novel"))
  )

p_drift_mid <- ggplot(
  pv_drift_discrim,
  aes(x = r, y = predictor, fill = condition, colour = condition)
) +
  stat_halfeye(alpha = 0.7, .width = c(0.5, 0.95)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "darkgrey") +
  scale_fill_manual(values = myColors_4) +
  scale_colour_manual(values = myColors_4) +
  labs(
    x = "Plausible r value",
    y = NULL,
    fill = NULL,
    colour = NULL
  ) +
  theme(
    legend.position = "bottom",
    text = element_text(size = 7, face = "bold"),
    strip.text = element_text(size = 7, face = "bold"),
    axis.text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12, face = "bold")
  )
```

```{r}
#| label: fig-plausible-corr-thresh
#| fig-cap: "Plausible correlation values distributions between threshold by response and recognition abilities. Points denote the median of the distribution and error bars represent 95% quantile intervals."
#| warning: false
#| fig-width: 10
#| fig-height: 5
p_thresh_left
```

```{r}
#| label: fig-plausible-corr-drift
#| fig-cap: "Plausible correlation values distributions between drift rate parameters by stimulus and recognition abilities. Points denote the median of the distribution and error bars represent 95% quantile intervals."
#| fig-subcap: 
#|  - "Plausible value correlations between drift rate (TRUE) and drift rate (FALSE) and recognition abilities for both stimulus conditions"
#|  - "Plausible value correlations between drift rate (TRUE-FASLE) and recognition abilities for both stimulus conditions"
#| layout-nrow: 2
#| warning: false
#| fig-width: 10
#| fig-height: 4
p_drift_first
p_drift_mid
```


```{r}
#| label: tbl-plausible-cor
#| tbl-cap: "Median r and 95% quantile intervals of the plausible value correlations between LBA parameters and recogntion ability scores."
#| include: true
#| tbl-height: 10
#| tbl-width: 15
table_data <- pv_summary |>
  filter(!is.na(label)) |>
  # no longer drop true/false rows
  mutate(
    r_fmt = sprintf("%.2f", r),
    ci_fmt = sprintf("[%.2f, %.2f]", .lower, .upper),
    color = case_when(
      .lower > 0 & .upper > 0 ~ "blue",
      .lower < 0 & .upper < 0 ~ "red",
      TRUE ~ "black"
    )
  ) |>
  select(label, predictor, r_fmt, ci_fmt, color) |>
  pivot_wider(
    names_from = predictor,
    values_from = c(r_fmt, ci_fmt, color),
    names_glue = "{predictor}_{.value}"
  ) |>
  separate(
    label,
    into = c("Parameter", "Condition"),
    sep = ": ",
    fill = "right"
  ) |>
  mutate(
    # extract accuracy before cleaning condition
    Accuracy = case_when(
      grepl("\\(TRUE\\)$", Condition, ignore.case = TRUE) ~ "TRUE",
      grepl("\\(FALSE\\)$", Condition, ignore.case = TRUE) ~ "FALSE",
      grepl("TRUE-FALSE", Condition, ignore.case = TRUE) ~ "TRUE-FALSE",
      TRUE ~ "-"
    ),
    # clean up condition labels
    Condition = case_when(
      Condition == "Learned (TRUE)" ~ "Learned",
      Condition == "Learned (FALSE)" ~ "Learned",
      Condition == "Novel (TRUE)" ~ "Novel",
      Condition == "Novel (FALSE)" ~ "Novel",
      Condition == "Learned (TRUE-FALSE)" ~ "Learned",
      Condition == "Novel (TRUE-FALSE)" ~ "Novel",
      TRUE ~ Condition
    ),
    # set ordering
    Parameter = factor(Parameter, levels = c("Threshold", "Drift")),
    Condition = factor(
      Condition,
      levels = c("Learned", "Novel")
    ),
    Accuracy = factor(Accuracy, levels = c("TRUE", "FALSE", "TRUE-FALSE", "-"))
  ) |>
  arrange(Parameter, Condition, Accuracy)


# rename columns to clean display names first
table_data_tt <- table_data |>
  rename(
    `median r_Face` = Face_r_fmt,
    `95% QI_Face` = Face_ci_fmt,
    `median r_Car` = Car_r_fmt,
    `95% QI_Car` = Car_ci_fmt
  ) |>
  select(
    Parameter,
    Condition,
    Accuracy,
    `median r_Face`,
    `95% QI_Face`,
    `median r_Car`,
    `95% QI_Car`,
    Face_color,
    Car_color
  )
table_data_tt <- as.data.frame(table_data_tt)

tt_table <- tt(
  table_data_tt |> select(-Face_color, -Car_color),
  notes = "Note. Blue = 95% QI entirely above zero; red = 95% QI entirely below zero."
) |>
  group_tt(j = list("Face" = 4:5, "Car" = 6:7)) |>
  style_tt(i = 0, bold = TRUE) |>
  style_tt(i = 1:nrow(table_data_tt), fontsize = 0.7) |>
  style_tt(i = 0, fontsize = 0.8) |>
  style_tt(i = -1, bold = TRUE, fontsize = 0.8) |>
  style_tt(
    i = which(table_data_tt$Face_color == "blue"),
    j = 4:5,
    color = "blue"
  ) |>
  style_tt(
    i = which(table_data_tt$Face_color == "red"),
    j = 4:5,
    color = "red"
  ) |>
  style_tt(
    i = which(table_data_tt$Car_color == "blue"),
    j = 6:7,
    color = "blue"
  ) |>
  style_tt(
    i = which(table_data_tt$Car_color == "red"),
    j = 6:7,
    color = "red"
  )

tt_table
```


## Bayesian linear regression models for threshold parameter

We ran two Bayesian linear regression models to estimate if face recognition ability could predict the threshold parameter (response cautions) for both the Learned and Novel responses. The models description can be found in @tbl-regression-model-tresh-formula. 

```{r}
#| label: regression-model-formula-thresh-face
#| include: false
# create a df
regression_formula <- tribble(
  ~Parameter          , ~Condition             , ~Ability     , ~Formula                                                                                     ,
  "Median threshold" , "response" , "Face"       , "Median threshold:  0 + response + face_ability + (1|participant)"              ,
  "Median drift rate" , "response" , "Face + Car" , "Median drift rate: 0 + response + face_ability + car_ability + (1|participant)" ,
)
```

```{r}
#| label: tbl-regression-model-tresh-formula
#| tbl-cap: "Formula specification for the pre-registered linear regression models."
#| include: true
regression_formula |>
  tt(
    notes = "response = two-choices response (Learned vs. Novel)"
  )
```

```{r}
#| label: read-thresh-models
#| include: false
b1_thresh <- readRDS(here("models/b1_thresh_face.rds"))
b2_thresh <- readRDS(here("models/b2_thresh_car.rds"))
```

```{r}
#| label: ppreds-thresh-wrangle
#| include: false
# --- all newdata grids ---

# b1: face only
newdata_b1 <- expand_grid(
  response = factor(
    c("learned", "novel"),
    levels = c("learned", "novel")
  ),
  face = seq(-15, 15, by = 1)
)

# b2: face + car at three car levels
newdata_b2 <- expand_grid(
  response = factor(
    c("learned", "novel"),
    levels = c("learned", "novel")
  ),
  face = seq(-15, 15, by = 1),
  car = c(-15, 0, 15)
)

# --- get posterior preds ---
preds_b1 <- newdata_b1 |>
  add_epred_draws(b1_thresh, ndraws = 1000, re_formula = NA) |>
  mutate(
    response = if_else(
      str_detect(as.character(response), "novel"),
      "Novel",
      "Learned"),
    response = factor(response, levels = c("Learned", "Novel")),
    model = "Face only"
  )

preds_b2 <- newdata_b2 |>
  add_epred_draws(b2_thresh, ndraws = 1000, re_formula = NA) |>
  mutate(
    response = if_else(
      str_detect(as.character(response), "novel"),
      "Novel",
      "Learned"
    ),
    response = factor(response, levels = c("Learned", "Novel")),
    model = "Face + Car",
    car_label = factor(
      car,
      levels = c(-15, 0, 15),
      labels = c("Low car (-15)", "Average car (0)", "High car (+15)")
    )
  )

# filter b2 at car = 0 only (for direct comparison with b1)
preds_b2_car0 <- preds_b2 |> filter(car == 0)

# shared y range across all predictions
y_range_thresh <- range(c(preds_b1$.epred, preds_b2$.epred))

##Face model
diff_summary_face <- preds_b1 %>%
  ungroup() %>%
  select(response, face, .draw, .epred) |>
  group_by(response, face) %>%
  median_qi(.epred, .width = 0.95)

##Face and Car model
# split condition back into learned/novel and true/false, then difference
diff_summary_car <- preds_b2_car0 %>%
  ungroup() %>%
  select(response, face, .draw, .epred) |>
  group_by(response, face) %>%
  median_qi(.epred, .width = 0.95)
```

```{r}
#| label: ppreds-thresh-plot-2x2
#| include: false
plot_preds_contrast <- function(diff_summary, title) {
  ggplot(
    diff_summary,
    aes(
      x = face,
      y = .epred,
      ymin = .lower,
      ymax = .upper,
      fill = response,
      colour = response
    )
  ) +
    geom_ribbon(alpha = 0.3, color = NA) +
    geom_line(linewidth = 1) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    facet_wrap(~response) +
    labs(
      title = title,
      x = "Face recognition (centred)",
      y = "Predicted threshold"
    ) +
    theme(legend.position = "none")
}

p_contrast_b1 <- plot_preds_contrast(diff_summary_face, "a) Face model")
p_contrast_b2 <- plot_preds_contrast(diff_summary_car, "b) Face and Car model")

p_contrast_b1_b2 <- (p_contrast_b1 | p_contrast_b2) +
  plot_layout(axis_titles = "collect", guides = "collect")
```

```{r}
get_slopes <- function(draws) {
  draws |>
    ungroup() |>
    filter(face %in% c(-15, 15)) |>
    group_by(.draw, response, across(any_of("car"))) |>
    summarise(
      slope = (.epred[face == 15] - .epred[face == -15]) / 30,
      .groups = "drop"
    )
}

slopes <- bind_rows(
  "Face"       = get_slopes(preds_b1),
  "Face + Car" = get_slopes(preds_b2_car0),
  .id = "Model"
)

thr_tab <- slopes |>
  group_by(Model, Response = response) |>
  median_qi(slope, .width = 0.95) |>
  ungroup() |>
  transmute(
    Parameter   = "Threshold slope",
    Model,
    Response,
    Estimate    = slope,
    `l-95% CI`  = .lower,
    `u-95% CI`  = .upper
  )


```


```{r}
#| label: fig-prediction-thresh
#| fig-cap: "Predicted threshold values from face recognition ability scores for a) Face model and b) Face and Car model "
#| layout-nrow: 1
#| warning: false
#| fig-width: 10
#| fig-height: 7
p_contrast_b1_b2
```

```{r}
#| label: tbl-pred-acc
#| tbl-cap: "Estimates and 95% quantile intervals of the predicted threshold values slope from face recognition ability scores"
#| include: true
#| tbl-height: 10
#| tbl-width: 15
tt(thr_tab) |>
  format_tt(j = 4:6, digits = 4, num_fmt = "decimal") |>
  style_tt(i = 0, bold = TRUE) |>
  style_tt(i = 1:nrow(thr_tab), fontsize = 0.7) |>
  style_tt(i = 0, fontsize = 0.8)
```


## Bayesian linear regression models: Car predictor

As an exploratory analysis, we were interested to see how does object recognition ability predict the drift rate and threshold values from the face recognition task. 
We ran two Bayesian linear regression models: the first one to estimate if object recognition ability could predict the response caution for learned and novel response and the second one to estimate if object recognition ability could predict the different drift rates (TRUE, FALSE and TRUE-FALSE) for learned and novel face stimuli. The models specificity can be found in @tbl-regression-model-formula-car. 

{{< pagebreak >}}

```{r}
#| label: regression-model-formula-car
#| include: false
# create a df
regression_formula <- tribble(
  ~Parameter          , ~Condition             , ~Ability     , ~Formula                                                                                     ,
  "Median threshold" , "response" , "Car"       , "Median threshold:  0 + response + car_ability + (1|participant)"              ,
  "Median drift rate" , "familiarity:accuracy" , "Car" , "Median drift rate: 0 + familiarity:accuracy + car_ability + (1|participant)" ,
)
```

```{r}
#| label: tbl-regression-model-formula-car
#| tbl-cap: "Formula specification for the pre-registered linear regression models."
#| include: true
regression_formula |>
  tt(
    notes = "response = two-choices response (Learned vs. Novel); familiarity = face stimulus familiarity (Learned vs. Novel); accuracy = response accuracy (TRUE vs. FALSE)"
  )
```

```{r}
#| label: read-models-car
#| include: false
b3_thresh <- readRDS(here("models/b3_thresh_car_only.rds"))
b3_drift <- readRDS(here("models/b3_drift_car_only.rds"))
```

```{r}
#| label: ppreds-thresh-car-wrangle
#| include: false
# --- all newdata grids ---

# b3: car only
newdata_b3 <- expand_grid(
  response = factor(
    c("learned", "novel"),
    levels = c("learned", "novel")
  ),
  car = seq(-15, 15, by = 1)
)

# --- get posterior preds ---
preds_b3 <- newdata_b3 |>
  add_epred_draws(b3_thresh, ndraws = 1000, re_formula = NA) |>
  mutate(
    response = if_else(
      str_detect(as.character(response), "novel"),
      "Novel",
      "Learned"),
    response = factor(response, levels = c("Learned", "Novel")),
    model = "Car only"
  )

##Car model
diff_summary_car <- preds_b3 %>%
  ungroup() %>%
  select(response, car, .draw, .epred) |>
  group_by(response, car) %>%
  median_qi(.epred, .width = 0.95)
```

*Threshold*: Regarding the threshold (response caution) predictions, visual inspections of the predictions show that as car recognition ability increases, the response caution for both, the Learned and Novel responses, increases (@fig-prediction-thresh-car).

```{r}
#| label: ppreds-thresh-car-plot-2x2
#| include: false
plot_preds_contrast <- function(diff_summary, title) {
  ggplot(
    diff_summary,
    aes(
      x = car,
      y = .epred,
      ymin = .lower,
      ymax = .upper,
      fill = response,
      colour = response
    )
  ) +
    geom_ribbon(alpha = 0.3, color = NA) +
    geom_line(linewidth = 1) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    facet_wrap(~response) +
    labs(
      title = title,
      x = "Car recognition (centred)",
      y = "Predicted threshold"
    ) +
    theme(legend.position = "none")
}

p_contrast_b3 <- plot_preds_contrast(diff_summary_car, "")
```

```{r}
#| label: fig-prediction-thresh-car
#| fig-cap: "Predicted threshold values from car recognition ability scores for learned and novel responses"
#| layout-nrow: 1
#| warning: false
#| fig-width: 10
#| fig-height: 7
p_contrast_b3
```

```{r}
#| label: ppreds-drift-car-wrangle
#| include: false
# --- all newdata grids ---

# b3: car only - drfit rate
newdata_b3 <- expand_grid(
  condition = factor(
    c("learned.false", "learned.true", "novel.false", "novel.true"),
    levels = c("learned.false", "learned.true", "novel.false", "novel.true")
  ),
  car = seq(-15, 15, by = 1)
)

# --- get posterior preds ---
preds_b3 <- newdata_b3 |>
  add_epred_draws(b3_drift, ndraws = 1000, re_formula = NA) |>
  mutate(
    stimulus = if_else(
      str_detect(as.character(condition), "novel"),
      "novel",
      "learned"
    ),
    accuracy = if_else(
      str_detect(as.character(condition), "true"),
      "TRUE",
      "FALSE"
    ),
    stimulus = factor(stimulus, levels = c("learned", "novel")),
    accuracy = factor(accuracy, levels = c("FALSE", "TRUE")),
    model = "Car only"
  )

# --- contrast data ---
contrasts_drift <- bind_rows(
  preds_b3 |>
    filter(car %in% c(-15, 15)) |>
    ungroup() |>
    mutate(car_level = if_else(car == 15, "high", "low"))
) |>
  select(model, condition, stimulus, accuracy, .draw, .epred, car_level) |>
  pivot_wider(names_from = car_level, values_from = .epred) |>
  mutate(
    diff = high - low
  )

# discrimination (true - false) per condition at high vs low face
discrim_contrast <- bind_rows(
  preds_b3 |>
    filter(car %in% c(-15, 15)) |>
    ungroup() |>
    mutate(car_level = if_else(car == 15, "high", "low"), car = 0)
) |>
  select(model, stimulus, accuracy, .draw, .epred, car_level) |>
  # create a combined key so we pivot both dimensions at once
  mutate(key = paste(accuracy, car_level, sep = "_")) |>
  select(-accuracy, -car_level) |>
  pivot_wider(names_from = key, values_from = .epred) |>
  mutate(
    discrim_high = TRUE_high - FALSE_high,
    discrim_low = TRUE_low - FALSE_low,
    discrim_diff = discrim_high - discrim_low
  )

# novel - learned discrimination contrast
# i.e., does face ability affect discrimination more for novel than learned?
learned_novel_contrast <- discrim_contrast |>
  select(model, stimulus, .draw, discrim_diff) |>
  pivot_wider(names_from = stimulus, values_from = discrim_diff) |>
  mutate(novel_vs_learned = novel - learned)
```

*Drift rate (TRUE and FALSE)*: Regarding the drift rate (TRUE) and (FALSE), visual inspections of the predictions show that, as car recognition ability increases, both drift rate (TRUE and FALSE) for both stimuli conditions (Learned and Novel) increases as well. In summary, as car recognition ability increases, the accumulation of signal (drift rate TRUE) and the accumulation of noise (FALSE) increase, regardless of the stimulus condition (@fig-prediction-drift-car). 

```{r}
#| label: ppreds-drift-plot-2x2
#| include: false
condition <- c("novel" = "Novel", "learned" = "Learned")
plot_preds_2x2 <- function(preds, title) {
  ggplot(
    preds,
    aes(
      x = car,
      y = .epred,
      fill = accuracy,
      colour = accuracy
    )
  ) +
    stat_lineribbon(alpha = 0.4, .width = c(0.66, 0.95)) +
    scale_fill_manual(values = myColors_acc_2) +
    scale_colour_manual(values = myColors_acc_2) +
    facet_wrap(~stimulus, nrow = 1, labeller = labeller(stimulus = condition)) +
    labs(
      title = title,
      x = "Car recognition (centred)",
      y = "Predicted drift rate",
      fill = NULL,
      colour = NULL
    ) +
    theme(legend.position = "bottom")
}

p_2x2_b3 <- plot_preds_2x2(preds_b3, "")
```

```{r}
#| label: fig-prediction-drift-car
#| fig-cap: "Predicted drfit rate (TRUE and FALSE) values from car recognition ability scores for learned and novel faces"
#| layout-nrow: 1
#| warning: false
#| fig-width: 10
#| fig-height: 7
p_2x2_b3
```

```{r}
#| label: ppreds-drift-constrasts
#| include: false
##Face model
# split condition back into learned/novel and true/false, then difference

diff_draws_car <- preds_b3 %>%
  ungroup() %>%
  separate(condition, into = c("learned_novel", "TRUE_FALSE"), sep = "\\.") %>%
  select(learned_novel, TRUE_FALSE, car, .draw, .epred) %>%
  pivot_wider(names_from = TRUE_FALSE, values_from = .epred) %>%
  mutate(diff = true - false)

# summarize with 95% QI, as a function of face
diff_summary_car <- diff_draws_car %>%
  group_by(learned_novel, car) %>%
  median_qi(diff, .width = 0.95)

#calculate the slopes
slope_draws_car <- b3_drift %>%
  spread_draws(
    `b_conditionlearned.true:car`,
    `b_conditionlearned.false:car`,
    `b_conditionnovel.true:car`,
    `b_conditionnovel.false:car`
  ) %>%
  mutate(
    learned_slope_diff = `b_conditionlearned.true:car` -
      `b_conditionlearned.false:car`,
    novel_slope_diff = `b_conditionnovel.true:car` -
      `b_conditionnovel.false:car`
  ) %>%
  select(.draw, learned_slope_diff, novel_slope_diff) %>%
  pivot_longer(-.draw, names_to = "learned_novel", values_to = "slope_diff")

slope_table_car <- slope_draws_car %>%
  group_by(learned_novel) %>%
  median_qi(slope_diff, .width = 0.95)
```

*Drift rate (TRUE-FALSE)*: Regarding the drift rate (TRUE-FALSE), or the signal-to-noise ratio provided by the stimulus, visual inspections of the predictions show that, as car recognition ability increases, only the drift rate for Novel faces increases, whereas the drift rate for Learned face does not follow this pattern. In summary, as car recognition ability increases, the ability to discriminate Novel faces increases but not the ability to discriminate Learned faces (@fig-prediction-drift-constrast-car).

```{r}
#| label: ppreds-drift-constrasts-plot-car
#| include: false
condition <- c("novel" = "Novel", "learned" = "Learned")
plot_preds_contrast <- function(diff_summary, title) {
  ggplot(
    diff_summary,
    aes(
      x = car,
      y = diff,
      ymin = .lower,
      ymax = .upper,
      fill = learned_novel,
      colour = learned_novel
    )
  ) +
    geom_ribbon(alpha = 0.3, color = NA) +
    geom_line(linewidth = 1) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    facet_wrap(~learned_novel, labeller = labeller(learned_novel = condition)) +
    labs(
      title = title,
      x = "Car recognition (centred)",
      y = "Predicted drift rate (TRUE − FALSE)"
    ) +
    theme(legend.position = "none")
}

p_contrast_b1 <- plot_preds_contrast(diff_summary_car, "")
```

```{r}
#| label: fig-prediction-drift-constrast-car
#| fig-cap: "Predicted drift rate (TRUE-FALSE) values from car recognition ability for learned and novel faces"
#| layout-nrow: 1
#| warning: false
#| fig-width: 10
#| fig-height: 7
p_contrast_b1
```
