---
title: "Supplementary Materials"
subtitle: "Individual differences in the computational processes that support face recognition"
running-head: "Individual differences and decision models of face memory"
author:
  - name: Raphaël Fournier
    orcid: 0000-0002-5135-4066
    affiliation:
      - ref: 1
  - name: Paul Downing
    affiliation: 
      - ref: 2
  - name: Richard Ramsey
    email: rramsey@ethz.ch
    orcid: 0000-0002-0329-2112
    corresponding: true
    affiliation:
      - ref: 1
      - ref: 3
      - ref: 4
affiliations:
  - id: 1
    name: Neural Control of Movement Laboratory
    department: Department of Health Sciences and Technology, ETH Zurich
  - id: 2
    name: 
    department: School of Psychology and Sports Science, Bangor University
  - id: 3
    name: Social Brain Sciences Laboratory
    department: Department of Humanities, Social and Political Sciences, ETH Zurich
  - id: 4
    name: ETH Zürich
    department: Learning and Instruction Laboratory, Department of Humanities, Social and Political Sciences
keywords: [face perception, memory, familiarity, person knowledge, decision models, evidence accumulation modelling]
bibliography: bibliography.bib
csl: apa.csl
functions: [place]
format:
  preprint-typst:
    output-file: supplementary.pdf
    fontsize: 11pt
    mainfont: Libertinus Serif
editor_options: 
  chunk_output_type: console
knitr:
  opts_chunk:
    dev: "ragg_png"
editor: 
  markdown: 
    wrap: sentence
---
```{r}
#| label: setup
#| include: false

pkgs <- c(
  "tidyverse", "RColorBrewer", "patchwork", "paletteer", 
  "here", "EMC2", "tidybayes", "tinytable", "osfr", "ragg"
)
vapply(
  pkgs, library, logical(1),
  character.only = TRUE, logical.return = TRUE
)

knitr::opts_chunk$set(echo = FALSE)

theme_set(
  theme_bw() +
    theme(text = element_text(size = 15, face = "bold"), 
          title = element_text(size = 15, face = "bold"),
          legend.position = "none")
)
## plot settings ##
## Plot colors
safe_pal <- paletteer_d("rcartocolor::Safe", 4)

myColors_3 <- c(
  "novel"   = safe_pal[1],
  "learned" = safe_pal[2]
)

# Subsets for plots if still needed
myColors_1 <- myColors_3[c("novel", "learned")]

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

{{< pagebreak >}}

## Parameter recovery study



{{< pagebreak >}}

## Priors

These are the priors used for the preregistered LBA model.

```{r}
prior_lba_table <- tibble(
  parameter = c(
    "v_SNovel:lMFALSE", "v_SLearned:lMFALSE",
    "v_SNovel:lMTRUE",  "v_SLearned:lMTRUE",
    "sv_lMTRUE",
    "B_lRNovel", "B_lRLearned",
    "A", "t0"
  ),
  description = c(
    "Drift rate", "Drift rate",
    "Drift rate", "Drift rate",
    "SD drift rate (true accumulator)",
    "Threshold", "Threshold",
    "Start point", "Non-decision time"
  ),
  condition = c(
    "Novel", "Learned",
    "Novel", "Learned",
    "-",
    "-", "-",
    "-", "-"
  ),
  correct = c(
    "False", "False",
    "True",  "True",
    "True",
    "-", "-",
    "-", "-"
  ),
  response = c(
    "-", "-", "-", "-", "-",
    "Novel", "Learned",
    "-", "-"
  ),
  mean = c(
    "0", "0", "1", "1",
    "log(0.5)",
    "log(1)", "log(1)", "log(0.5)", "log(0.3)"
  ),
  sd = c(
    "2", "2", "2", "2",
    "0.5", "0.5", "0.5", "0.3", "0.3"
  )
)

## create the table
prior_lba_table |> 
  tt() |> 
  style_tt(fontsize = 0.85)
```

## Model checks

### Convergence

The chains showed satisfactory mixing as evidenced by the rhat values.

```{r}
#| label: fig-chains-exp1
#| fig-cap: "MCMC chains"
#| fig-width: 8
#| fig-height: 6


```

### Posterior predictive checks

```{r}
#| label: fig-ppreds-exp1
#| fig-cap: "Posterior predictive checks"
#| fig-subcap:
#|   - "CDF"
#|   - "Density"
#| layout-nrow: 2
#| out-height: "40%"

```

### Prior vs posterior plots

```{r}
#| label: fig-prior-post-exp1
#| fig-cap: "Priors vs posterior plots"
#| fig-width: 8
#| fig-height: 6

```


{{< pagebreak >}}

## Point estimate correlations with Cambridge Car Memory scores

```{r}

```


## Plausible value correlations with Cambridge Car Memory scores

```{r}

```


## Bayesian linear regression models for threshold parameter

```{r}

```



