---
title: "exploratory_analysis"
author: "Fournier Raphaël"
format: html
editor: source
editor_options: 
  chunk_output_type: console
---

This file calculates, visualises and summarises the exploratory analyses from the individual differences face familiarity study.

```{r}
#| label: setup
#| include: false

pkgs <- c(
  "EMC2", "tidyverse", "tidybayes", "patchwork", "paletteer", "here",
  "corrplot", "ggdist", "RColorBrewer", "parallel", "future", "brms", "tidyr",
  "ggplot2", "dplyr", "ggpubr"
  )
vapply(
  pkgs, library, logical(1), character.only = TRUE, logical.return = TRUE
  )

theme_set(
  theme_bw() +
    theme(
      text  = element_text(size = 15, face = "bold"),
      title = element_text(size = 15, face = "bold"),
      legend.position = "none"
    )
)

## set options
options(mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()

detectCores()

plan(multicore)

## make things reproducible where possible 
set.seed(123)

safe_pal   <- paletteer_d("rcartocolor::Safe", 4)
dark_pal <- paletteer_d("RColorBrewer::Dark2", 2)

mycolors <- c("novel" = safe_pal[1], "learned" = safe_pal[2])
mycolors_acc <- c(
  "TRUE" = dark_pal[1],
  "FALSE" = dark_pal[2]
)
```


## Correlation between drift rate and d'
1. Read in the data

```{r}
#| label: read-wrangle

# read in posterior medians and cambridge scores
theta_pid_q <- read_csv(
  here("data", "model", "theta_pid_q.csv"),
  show_col_types = FALSE
)
cam_face_car <- read_csv(
  here("data","model", "cambridge_acc_scores_pid_res.csv"))


# drift rate data — same structure as pilot
drift_data <- theta_pid_q |>
  filter(str_detect(param, "v_S")) |>
  select(pid, param, value) |>
  mutate(
    stimulus = if_else(str_detect(param, "novel"), "novel", "learned"),
    accuracy = if_else(str_detect(param, "TRUE"), "true", "false"),
    condition = factor(
      paste0(stimulus, ".", accuracy),
      levels = c("learned.false", "novel.false", "learned.true", "novel.true")
    ),
    stimulus = factor(stimulus, levels = c("learned", "novel")),
    accuracy = factor(accuracy, levels = c("false", "true"))
  ) |>
  left_join(cam_face_car, by = "pid") |>
  select(pid, condition, stimulus, accuracy, face, car, value)

# threshold data
thresh_data <- theta_pid_q |>
  filter(str_detect(param, "B_")) |>
  select(pid, param, value) |>
  mutate(
    response = if_else(str_detect(param, "novel"), "novel", "learned"),
    response = factor(response, levels = c("learned", "novel"))
  ) |>
  left_join(cam_face_car, by = "pid") |>
  select(pid, response, face, car, value)

# check structure
head(drift_data)
head(thresh_data)
```

#Calculate d' and Criterion from the accuracy data of the face recognition taks

```{r}
#read in the face recognition task
data_task <- read_csv(here("data", "data.csv"))

#create d' dataframes: separate data frames for learned and novel (we will put them back together later)
dprime_final <- d_prime %>%
  group_by(pid) %>%
  summarise(
    # Learned condition
    HR_learned = mean(resp_type_learned == "hit", na.rm = TRUE),
    FAR_learned = mean(resp_type_learned == "fa",  na.rm = TRUE),
    
    # Novel condition
    HR_novel = mean(resp_type_novel == "hit", na.rm = TRUE),
    FAR_novel = mean(resp_type_novel == "fa",  na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    dprime_learned = qnorm(HR_learned) - qnorm(FAR_learned),
    dprime_novel   = qnorm(HR_novel)   - qnorm(FAR_novel)
  )
```

#Adding d'prime column in drift dataframe
```{r}
cor_data <- theta_pid_q |>
  select(pid, param, value) |>
  pivot_wider(names_from = param, values_from = value) |>
  left_join(dprime_final, by = "pid") |>
  # derived contrasts
  mutate(
    # drift discrimination (true - false) per condition
    discrim_novel   = v_Snovel_lMTRUE   - v_Snovel_lMFALSE,
    discrim_learned = v_Slearned_lMTRUE - v_Slearned_lMFALSE,
    # cross-condition drift contrast (novel - learned discrimination)
    discrim_diff    = discrim_novel - discrim_learned,
    # threshold contrast
    thresh_diff     = B_lRnovel - B_lRlearned
  )
```


#Correlations
```{r}
param_cols <- c(
  "drift_learned", "drift_novel"
)

# pivot to long with both predictors as a variable
cor_long <- cor_data |>
  select(pid, discrim_novel, discrim_learned, dprime_learned, dprime_novel)|>
  rename(drift_learned = "discrim_learned", drift_novel = "discrim_novel") |>
  pivot_longer(
    cols      = all_of(param_cols),
    names_to  = "param",
    values_to = "estimate"
  ) |>
  pivot_longer(
    cols      = c(dprime_learned, dprime_novel),
    names_to  = "predictor",
    values_to = "score"
  ) |>
  mutate(
    predictor = if_else(str_detect(predictor, "learned"), "sensitivity learned", "sensitivity novel"),
    predictor = factor(predictor, levels = c("sensitivity learned", "sensitivity novel"))
  )

# compute r per param x predictor
cor_labels <- cor_long |>
  group_by(param, predictor) |>
  summarise(
    r = cor(estimate, score, use = "complete.obs"),
    .groups = "drop"
  ) |>
  mutate(
    label = sprintf("r = %.2f", r),
    vjust = if_else(predictor == "sensitivity learned", 1.5, 3.0)
  )

# Extract the two panels' data
df_plot <- cor_long %>%
  # Keep only the matching condition pairs
  filter(
    (param == "drift_learned" & predictor == "sensitivity learned") |
    (param == "drift_novel"   & predictor == "sensitivity novel")
  ) %>%
  mutate(condition = case_when(
    param == "drift_learned" ~ "Learned",
    param == "drift_novel"   ~ "Novel"
  ))

# Plot
ggplot(df_plot, aes(x = score, y = estimate)) +
  geom_point(size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, color = "steelblue") +
  stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "top") + # requires ggpubr
  facet_wrap(~ condition, scales = "free") +
  labs(
    x = "Sensitivity (d')",
    y = "Drift rate",
    title = "Drift rate ~ Sensitivity by condition"
  ) +
  theme_bw()

ggsave(here("figures", "exploratory", "d_prime.jpeg"), 
       width = 10, height = 8)
```



## Bayesian linear regression model: Fitting Car only models

### Fit model b3 threshold

```{r}
#| label: fit-b3-thresh

## formula
formula = bf(value ~ 0 + response + response:car + (1|pid))

## check priors
get_prior(formula, data = thresh_data, family = gaussian())

## set priors
priors <- c(
  set_prior("normal(2, 1)", class = "b"),
  set_prior("normal(0, 1)", class = "b", coef = c("responselearned:car",
                                                  "responsenovel:car")),
  set_prior("normal(0, 1)", class = "sd", group = "pid"),
  set_prior("normal(0, 1)", class = "sigma")
)

## fit the model
b3_thresh <- fit_model(
  formula = formula,
  prior = priors,
  data = thresh_data,
  file = "models/iid/b3_thresh_car_only",
)

summary(b3_thresh)
```

### fit model b3 drift rate

```{r}
#| label: fit-b3-drift

## formula
formula = bf(value ~ 0 + condition + condition:car + (1|pid))

## check priors
get_prior(formula, data = drift_data, family = gaussian())

## set priors
priors <- c(
  set_prior("normal(3, 1)", class = "b"),
  set_prior("normal(0, 1)", class = "b", coef = c("conditionlearned.false:car",
                                                  "conditionlearned.true:car",
                                                  "conditionnovel.false:car",
                                                  "conditionnovel.true:car")),
  set_prior("normal(0, 1)", class = "sd", group = "pid"),
  set_prior("normal(0, 1)", class = "sigma")
)

## fit the model
b3_drift <- fit_model(
  formula = formula,
  prior = priors,
  data = drift_data,
  file = "models/iid/b3_drift_car_only",
)

summary(b3_drift)
```

#Read models if already fitted
```{r}
b3_thresh <- readRDS(here("models", "iid", "b3_thresh_car_only.rds"))
b3_drift <- readRDS(here("models", "iid", "b3_drift_car_only.rds"))
```


#Plotting predictions

```{r}
#| label: ppreds-thresh-car-wrangle
#| include: false
# --- all newdata grids ---

# b3: car only
newdata_b3 <- expand_grid(
  response = factor(
    c("learned", "novel"),
    levels = c("learned", "novel")
  ),
  car = seq(-15, 15, by = 1)
)

# --- get posterior preds ---
preds_b3_thresh <- newdata_b3 |>
  add_epred_draws(b3_thresh, ndraws = 1000, re_formula = NA) |>
  mutate(
    response = if_else(
      str_detect(as.character(response), "novel"),
      "Novel",
      "Learned"),
    response = factor(response, levels = c("Learned", "Novel")),
    model = "Car only"
  )

##Car model
diff_summary_car_thresh <- preds_b3_thresh %>%
  ungroup() %>%
  select(response, car, .draw, .epred) |>
  group_by(response, car) %>%
  median_qi(.epred, .width = 0.95)
```

```{r}
#| label: ppreds-thresh-car-plot-2x2
#| include: false
plot_preds_contrast <- function(diff_summary, title) {
  ggplot(
    diff_summary,
    aes(
      x = car,
      y = .epred,
      ymin = .lower,
      ymax = .upper,
      fill = response,
      colour = response
    )
  ) +
    geom_ribbon(alpha = 0.3, color = NA) +
    geom_line(linewidth = 1) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    facet_wrap(~response) +
    labs(
      title = title,
      x = "Car recognition (centred)",
      y = "Predicted threshold"
    ) +
    theme(legend.position = "none")
}

p_contrast_b3 <- plot_preds_contrast(diff_summary_car_thresh, "")
p_contrast_b3
ggsave(here("figures", "exploratory", "preds_thresh_car_only.jpeg"))
```


```{r}
#| label: ppreds-drift-car-wrangle
# --- all newdata grids ---

# b3: car only - drfit rate
newdata_b3 <- expand_grid(
  condition = factor(
    c("learned.false", "learned.true", "novel.false", "novel.true"),
    levels = c("learned.false", "learned.true", "novel.false", "novel.true")
  ),
  car = seq(-15, 15, by = 1)
)

# --- get posterior preds ---
preds_b3_drift <- newdata_b3 |>
  add_epred_draws(b3_drift, ndraws = 1000, re_formula = NA) |>
  mutate(
    stimulus = if_else(
      str_detect(as.character(condition), "novel"),
      "novel",
      "learned"
    ),
    accuracy = if_else(
      str_detect(as.character(condition), "true"),
      "TRUE",
      "FALSE"
    ),
    stimulus = factor(stimulus, levels = c("learned", "novel")),
    accuracy = factor(accuracy, levels = c("FALSE", "TRUE")),
    model = "Car only"
  )

# --- contrast data ---
contrasts_drift <- bind_rows(
  preds_b3_drift |>
    filter(car %in% c(-15, 15)) |>
    ungroup() |>
    mutate(car_level = if_else(car == 15, "high", "low"))
) |>
  select(model, condition, stimulus, accuracy, .draw, .epred, car_level) |>
  pivot_wider(names_from = car_level, values_from = .epred) |>
  mutate(
    diff = high - low
  )

# discrimination (true - false) per condition at high vs low face
discrim_contrast <- bind_rows(
  preds_b3 |>
    filter(car %in% c(-15, 15)) |>
    ungroup() |>
    mutate(car_level = if_else(car == 15, "high", "low"), car = 0)
) |>
  select(model, stimulus, accuracy, .draw, .epred, car_level) |>
  # create a combined key so we pivot both dimensions at once
  mutate(key = paste(accuracy, car_level, sep = "_")) |>
  select(-accuracy, -car_level) |>
  pivot_wider(names_from = key, values_from = .epred) |>
  mutate(
    discrim_high = TRUE_high - FALSE_high,
    discrim_low = TRUE_low - FALSE_low,
    discrim_diff = discrim_high - discrim_low
  )

# novel - learned discrimination contrast
# i.e., does face ability affect discrimination more for novel than learned?
learned_novel_contrast <- discrim_contrast |>
  select(model, stimulus, .draw, discrim_diff) |>
  pivot_wider(names_from = stimulus, values_from = discrim_diff) |>
  mutate(novel_vs_learned = novel - learned)
```

```{r}
#| label: ppreds-drift-plot-2x2-supp
#| include: false
condition <- c("novel" = "Novel", "learned" = "Learned")
plot_preds_2x2 <- function(preds, title) {
  ggplot(
    preds,
    aes(
      x = car,
      y = .epred,
      fill = accuracy,
      colour = accuracy
    )
  ) +
    stat_lineribbon(alpha = 0.4, .width = c(0.66, 0.95)) +
    scale_fill_manual(values = mycolors_acc) +
    scale_colour_manual(values = mycolors_acc) +
    facet_wrap(~stimulus, nrow = 1, labeller = labeller(stimulus = condition)) +
    labs(
      title = title,
      x = "Car recognition (centred)",
      y = "Predicted drift rate",
      fill = NULL,
      colour = NULL
    ) +
    theme(legend.position = "bottom")
}

p_2x2_b3 <- plot_preds_2x2(preds_b3_drift, "")
p_2x2_b3
ggsave(here("figures", "exploratory", "preds_drift_true_false_car_only.jpeg"))
```

```{r}
#| label: ppreds-drift-constrasts-supp
#| include: false
##Face model
# split condition back into learned/novel and true/false, then difference

diff_draws_car <- preds_b3 %>%
  ungroup() %>%
  separate(condition, into = c("learned_novel", "TRUE_FALSE"), sep = "\\.") %>%
  select(learned_novel, TRUE_FALSE, car, .draw, .epred) %>%
  pivot_wider(names_from = TRUE_FALSE, values_from = .epred) %>%
  mutate(diff = true - false)

# summarize with 95% QI, as a function of face
diff_summary_car_drift <- diff_draws_car %>%
  group_by(learned_novel, car) %>%
  median_qi(diff, .width = 0.95)

#calculate the slopes
slope_draws_car <- b3_drift %>%
  spread_draws(
    `b_conditionlearned.true:car`,
    `b_conditionlearned.false:car`,
    `b_conditionnovel.true:car`,
    `b_conditionnovel.false:car`
  ) %>%
  mutate(
    learned_slope_diff = `b_conditionlearned.true:car` -
      `b_conditionlearned.false:car`,
    novel_slope_diff = `b_conditionnovel.true:car` -
      `b_conditionnovel.false:car`
  ) %>%
  select(.draw, learned_slope_diff, novel_slope_diff) %>%
  pivot_longer(-.draw, names_to = "learned_novel", values_to = "slope_diff")

slope_table_car <- slope_draws_car %>%
  group_by(learned_novel) %>%
  median_qi(slope_diff, .width = 0.95)
```

```{r}
#| label: ppreds-drift-constrasts-plot-car
#| include: false
condition <- c("novel" = "Novel", "learned" = "Learned")
plot_preds_contrast <- function(diff_summary, title) {
  ggplot(
    diff_summary,
    aes(
      x = car,
      y = diff,
      ymin = .lower,
      ymax = .upper,
      fill = learned_novel,
      colour = learned_novel
    )
  ) +
    geom_ribbon(alpha = 0.3, color = NA) +
    geom_line(linewidth = 1) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    facet_wrap(~learned_novel, labeller = labeller(learned_novel = condition)) +
    labs(
      title = title,
      x = "Car recognition (centred)",
      y = "Predicted drift rate (TRUE − FALSE)"
    ) +
    theme(legend.position = "none")
}

p_contrast_b1 <- plot_preds_contrast(diff_summary_car_drift, "")
p_contrast_b1
ggsave(here("figures", "exploratory", "preds_drift_contrasts_car_only.jpeg"))
```

